### 2009-06-08 Paper/paper.tex
starting paper

\documentclass{acmsiggraph}                     % final
%\documentclass[annualconference]{acmsiggraph}  % final (annual conference)
%\documentclass[review]{acmsiggraph}            % review
%\documentclass[widereview]{acmsiggraph}        % wide-spaced review
%\documentclass[preprint]{acmsiggraph}          % preprint

%% Uncomment one of the five lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and ``final'' is for
%% the version to be printed. The ``final'' variant will accept the 
%% ``annualconference'' parameter, which changes the height of the space
%% left clear for the ACM copyright information.

%% The 'helvet' and 'times' packages define the typefaces used for
%% serif and sans serif type in this document. Computer Modern Roman 
%% is used for mathematics typesetting. The scale factor is set to .92
%% to bring the sans-serif type in line with the serif type.

\usepackage[scaled=.92]{helvet}
\usepackage{times}

%% The 'graphicx' package allows for the inclusion of EPS figures.

\usepackage{graphicx}

%% use this for zero \parindent and non-zero \parskip, intelligently.

\usepackage{parskip}

%% Optional: the 'caption' package provides a nicer-looking replacement
%% for the standard caption environment. With 'labelfont=bf,'textfont=it',
%% caption labels are bold and caption text is italic.

\usepackage[labelfont=bf,textfont=it]{caption}

%% If you are submitting a paper to the annual conference, please replace 
%% the value ``0'' below with the numeric value of your OnlineID. 
%% If you are not submitting this paper to the annual conference, 
%% you may safely leave it at ``0'' -- it will not be included in the output.

\onlineid{0}

%% Paper title.

\title{Global Illumination for Fun and Profit}

%% Author and Affiliation (single author).

%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\Allied Widgets Research}

%% Author and Affiliation (multiple authors).

\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research}

%% Keywords that describe your work.

\keywords{radiosity, global illumination, constant time}

%%%%%% START OF THE PAPER %%%%%%

\begin{document}

%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

\maketitle

%% Abstract section.

\begin{abstract}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Citations can be done this way~\cite{Jobs95} or this more concise 
way~\shortcite{Jobs95}, depending upon the application.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
in vulputate velit esse molestie~\cite{notes2002} 
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}

\end{abstract}

%% ACM Computing Review (CR) categories. 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CRcat'' command takes four arguments.

\begin{CRcatlist}
  \CRcat{K.6.1}{Management of Computing and Information Systems}%
{Project and People Management}{Life Cycle};
  \CRcat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
\end{CRcatlist}

%% The ``\keywordlist'' command prints out the keywords.
\keywordlist

\section{Introduction}

%% The ``\copyrightspace'' command must be the first command after the 
%% start of the first section of the body of your paper. It ensures the
%% copyright space is left at the bottom of the first column on the first
%% page of your paper.

%% \copyrightspace

\section{Exposition}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\begin{equation}
 \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

\begin{eqnarray}
x & \ll & y_{1} + \cdots + y_{n} \\
  & \leq & z
\end{eqnarray}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

%\begin{figure*}[ht]
%\centering
%\includegraphics[width=2.5in]{sample.eps}
%\caption{Sample illustration.}
%\end{figure*}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\section{A New Section}

\subsection{A New Subsection}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

\subsection{A Second Subsection}

\subsubsection{A Subsubsection}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

\subsubsection{Another Subsubsection}

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

\subsection{A Final Subsection}

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

\section{Back to the Section}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

\section{Conclusion}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
invidunt ut labore et dolore magna aliquyam erat, sed diam
voluptua. At vero eos et accusam et justo duo dolores et ea
rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
molestie consequat, vel illum dolore eu feugiat nulla facilisis at
vero eros et accumsan et iusto odio dignissim qui blandit praesent
luptatum zzril delenit augue duis dolore te feugait nulla
facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
aliquam erat volutpat.

Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
vel eum iriure dolor in hendrerit in vulputate velit esse molestie
consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
delenit augue duis dolore te feugait nulla facilisi.

\section*{Acknowledgements}

To Robert, for all the bagels.

\bibliographystyle{acmsiggraph}
\nocite{*}
\bibliography{template}
\end{document}

### 2009-06-10 Results/script.R
working on script.R

@@ -1,0 +1,96 @@
+datafilename="all_results.csv"
+data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
+
+#data.all[0,1] = 
+#data.all[0,2] = 'fov'
+#data.all[0,3] = 'deadlen'
+data.all[,'mean'] = rowMeans(data.all[,4:120])
+
+#plot(data.all[,'avg'],type='lines')
+
+attach(data.all)
+data.agg = aggregate(mean,list(V1,V2,V3), mean)
+colnames(data.agg) <- c("id","fov","deadlen",'mean')
+
+data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
+data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
+data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
+
+data.agg[sapply(data.agg['deadlen'],function(x) x==1),'latency'] = 'lowlat'
+data.agg[sapply(data.agg['deadlen'],function(x) x==6),'latency'] = 'medlat'
+data.agg[sapply(data.agg['deadlen'],function(x) x==12),'latency'] = 'upplat'
+
+#data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
+#data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
+#data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
+#data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
+#data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
+#data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
+
+#data.agg[sapply(data.agg['game'],function(x) x==1),'game'] = '0-1'
+#data.agg[sapply(data.agg['game'],function(x) x==3),'game'] = '2-5'
+#data.agg[sapply(data.agg['game'],function(x) x==10),'game'] = '5-15'
+
+#data.agg[4] = log(data.agg[4])
+#data.agg[17,4] <- -2
+
+detach(data.all)
+
+#aov.gender=aov(recall~gender,data.agg)
+#print(summary(aov.gender))
+#boxplot(recall~gender,data=data.agg)
+
+#aov.game=aov(recall~game,data.agg)
+#print(summary(aov.game))
[…]

### 2009-06-10 trunk/Results/script.R
initial analysis

@@ -16,13 +16,14 @@ data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
 data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
 data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
 
-data.agg[sapply(data.agg['deadlen'],function(x) x==1),'latency'] = 'lowlat'
-data.agg[sapply(data.agg['deadlen'],function(x) x==6),'latency'] = 'medlat'
-data.agg[sapply(data.agg['deadlen'],function(x) x==12),'latency'] = 'upplat'
+data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
+data.agg[sapply(data.agg['deadlen'],function(x) x==1),'deadlen'] = 'medlen'
+data.agg[sapply(data.agg['deadlen'],function(x) x==2),'deadlen'] = 'upplen'
 
-#data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
-#data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
-#data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
+data.agg[sapply(data.agg['id'],function(x) x==0),'id'] = 'id0'
+data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
+data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
+data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
 #data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
 #data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
 #data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
@@ -45,24 +46,24 @@ detach(data.all)
 #boxplot(recall~game,data=data.agg)
 
 # one-way repeated measures anova on fov
-#aov.fov=aov(recall~fov+Error(id/fov),data.agg)
+#aov.fov=aov(mean~fov+Error(id/fov),data.agg)
 #print(summary(aov.fov))
-#boxplot(recall~fov,data=data.agg)
+#boxplot(mean~fov,data=data.agg)
 
 
-# one-way repeated measures anova on latency
[…]

### 2009-06-10 Paper/paper.tex
starting paper

@@
 
-\title{Global Illumination for Fun and Profit}
+\title{Title Goes Here}
 
@@
 
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
-\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research}
+\author{
+%Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
+%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
+%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research}
+}
 
@@
 
-\keywords{radiosity, global illumination, constant time}
+\keywords{
+%radiosity, global illumination, constant time
+}
 
@@
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at 
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Citations can be done this way~\cite{Jobs95} or this more concise 
-way~\shortcite{Jobs95}, depending upon the application.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit~\cite{Pellacini:2005:LAH}
-in vulputate velit esse molestie~\cite{notes2002} 
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.~\cite{Park:2006:DSI}
-
 \end{abstract}
@@
 \begin{CRcatlist}
-  \CRcat{K.6.1}{Management of Computing and Information Systems}%
-{Project and People Management}{Life Cycle};
-  \CRcat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
+%  \CRcat{K.6.1}{Management of Computing and Information Systems}{Project and People Management}{Life Cycle};
+%  \CRcat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
 \end{CRcatlist}
@@
 
-\section{Exposition}
+\section{Related Work}
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
+\cite{4637329,4811058,1383060}
 
-\begin{equation}
- \sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
+\section{Experimental Design}
 
-\begin{eqnarray}
-x & \ll & y_{1} + \cdots + y_{n} \\
-  & \leq & z
-\end{eqnarray}
+Our experiment involves a tracking task which uses an augmented reality interface to allow people to be tracked behind walls.
+A participant stands in the middle of a square room with doors and windows on all sides.
+Outside of the room, people are walking around the building.
+The target person to be tracked is visually distinguishable by a large black top hat.
+The augmented reality interface overlays a translucent red rectangle on each person, which is visible even when the person is occluded by a wall.
+However, the target person's overlay is not different from any other person's overlay; the target is only identifiable when visible through a window or a door.
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
+To actually run this experiment using a true augmented reality system would be difficult.
+Many confederates would be needed to walk around the room, and their movements would need to be accurately reproduced for each new participant.
+Also, the confederates would need to be continously tracked for the augmented overlays to be displayed.
+Instead, we simulated the experiment setup in a completely virtual environment.
 
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
+Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
+The tracking overlays are  augmented on top of the image to form the simulated augmented reality view.
 
-%\begin{figure*}[ht]
-%\centering
-%\includegraphics[width=2.5in]{sample.eps}
-%\caption{Sample illustration.}
-%\end{figure*}
+By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
+Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
+The parameters of our experiment are divided between the ``real'' immersion factors, such as the field of view of the HMD, and the ``augmented'' immersion factors, such as the field of view of the simulated AR display, and the performance of the head tracker.
+We introduce periods of tracking failure, where the augmented overlays disappear, to simulate the effect of failures in a real tracking system.
+For example, such failures might occur with a magnetic tracker near interfering materials, or with visual tracking when tracked features are lost.
 
-\section{A New Section}
+In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.
+Each trial lasted sixty seconds, and had seven tracking failures.
+The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
+We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).
 
-\subsection{A New Subsection}
+\section{Results}
 
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
+\section{Analysis}
 
-\subsection{A Second Subsection}
-
-\subsubsection{A Subsubsection}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-\subsubsection{Another Subsubsection}
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-\subsection{A Final Subsection}
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-\section{Back to the Section}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
-
-\section{Conclusion}
-
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-invidunt ut labore et dolore magna aliquyam erat, sed diam
-voluptua. At vero eos et accusam et justo duo dolores et ea
-rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-ipsum dolor sit amet.
-
-Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-vero eros et accumsan et iusto odio dignissim qui blandit praesent
-luptatum zzril delenit augue duis dolore te feugait nulla
-facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-aliquam erat volutpat.
-
-Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et
-accumsan et iusto odio dignissim qui blandit praesent luptatum zzril
-delenit augue duis dolore te feugait nulla facilisi.
+\section{Conclusions}
 
@@
 
-To Robert, for all the bagels.
-
 \bibliographystyle{acmsiggraph}
 \nocite{*}
-\bibliography{template}
+\bibliography{paper}
 \end{document}

### 2009-06-10 Code/path.py
fixed error recording to record every frame

@@ -133,9 +133,10 @@ class PathGenerator:
 		self.num_nearby += num_nearby
 		self.num_samples += 1
 		
-	def checkError(self,tophat,errlist):
+	def checkError(self,tophat):
 		err =  error.MeasureError(tophat)
-		errlist.append(err)
+		self.errlist.append(err)
+		self.timelist.append( viz.tick() - self.starttime )
 		#print "current error: ",err
 		#a = 0
 	
@@ -152,18 +153,19 @@ class PathGenerator:
 		print "here6"
 		self.num_nearby = 0
 		self.num_samples = 0
-		nearby_timer = vizact.ontimer(1/speedMultiplier,self.checkNearby,tophat,peopleset)
+		#nearby_timer = vizact.ontimer(1/speedMultiplier,self.checkNearby,tophat,peopleset)
 		
-		visible_timer = vizact.ontimer(0.5/speedMultiplier,tophat.checkVisibleTime)
+		#visible_timer = vizact.ontimer(0.5/speedMultiplier,tophat.checkVisibleTime)
 		
-		errlist = []
-		error_timer = vizact.ontimer(0.1/speedMultiplier,self.checkError,tophat,errlist)
+		#errlist = []
+		#error_timer = vizact.ontimer(0.1/speedMultiplier,self.checkError,tophat,errlist)
+		#error_timer = vizact.onupdate(-10,self.checkError,tophat,errlist,timelist)
 		print "here7"
 		yield viztask.waitTime(self.taskTime)
 		print "here8"
-		vizact.removeEvent(visible_timer)
-		vizact.removeEvent(nearby_timer)
-		vizact.removeEvent(error_timer)
+		#vizact.removeEvent(visible_timer)
+		#vizact.removeEvent(nearby_timer)
+		#vizact.removeEvent(error_timer)
 		
 		#save the path
 		print "peopleSetLength:",len(peopleset)
@@ -180,7 +182,7 @@ class PathGenerator:
 		ps.collisions = ps.abePath.collisions
 		ps.timeNotVisible = tophat.timeNotVisible*speedMultiplier
 		ps.pointsWalkedTo = len( ps.abePath.points )
-		ps.numberPeopleNearby = self.num_nearby / self.num_samples
+#		ps.numberPeopleNearby = self.num_nearby / self.num_samples
 		ps.quadrantsReached = ps.abePath.quadrantsReached
 		print "num collisions: ",ps.collisions
[…]

### 2009-06-10 Code/run_experiment.py

@@ -98,6 +98,10 @@ def run_tasks():
 		for err in pg.errlist:
 		    resultsfile.write("%f, "%err)
 		resultsfile.write('\n')
+		resultsfile.write(",,,")
+		for time in pg.timelist:
+		    resultsfile.write("%f, "%time)
+		resultsfile.write('\n')
 	
 		# close results file
 		# using close command

### 2009-06-10 trunk/NewResults/resampler.py
adding resampler and R script

@@ -1,0 +1,79 @@
+import sys
+
+class Resampler:
+    def __init__(self,samples,timing,rate):
+	self.timestep = 1. / rate
+	self.samples = samples
+	self.timing = timing
+	self.samples.insert(0,samples[0])
+	self.timing.insert(0,0)
+	lastsample = self.samples[len(self.samples)-1]
+	lasttiming = self.timing[len(self.timing)-1]
+	for i in xrange(100):
+	    self.samples.append(lastsample)
+	    lasttiming += self.timestep
+	    self.timing.append(lasttiming)
+	self.time = 0
+    
+    def step(self):
+	for i in xrange( len(self.samples)-1 ):
+	    if self.time >= self.timing[i] and self.time <= self.timing[i+1]:
+		t0 = self.timing[i]
+		t1 = self.timing[i+1]
+		s0 = self.samples[i]
+		s1 = self.samples[i+1]
+		alpha = ( self.time - t0 ) / ( t1 - t0 )
+		break
+	sample = alpha * s0 + ( 1. - alpha ) * s1
+	self.time += self.timestep
+	return (self.time,sample)
+
+for i in xrange(27):
+    samples_line = sys.stdin.readline()
+    timing_line = sys.stdin.readline()
+    list = samples_line.rstrip().split(',')
+    id = list[0]
+    fov = list[1]
+    deadlen = list[2]
+    samples = list[3:]
+    list = timing_line.rstrip().split(',')
+    timing = list[3:]
+
+    samples_in = []
+    timing_in = []
+    for j in samples:
+	try:
+	    samples_in.append( float(j) )
+	except:
+	    continue
+    for j in timing:
+	try:
+	    timing_in.append( float(j) )
+	except:
+	    continue
+
+    #print "%d samples, %d timings"%(len(samples_in),len(timing_in))
+    rs = Resampler( samples_in, timing_in, 60. )
+
+    samples_out = []
+    timing_out = []
+
+    sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+    for j in xrange(60*60):
+	(time,sample) = rs.step()
+	sys.stdout.write("%f"%sample)
+	if j != 3599:
+	    sys.stdout.write(",")
+	#samples_out.append(sample)
+	#timing_out.append(time)
+    sys.stdout.write("\n")
+    
+
+
+
+
+
+
+
+
+    

### 2009-06-10 trunk/NewResults/script.R

@@ -1,0 +1,89 @@
+datafilename="resampled.csv"
+data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
+
+#data.all[0,1] = 
+#data.all[0,2] = 'fov'
+#data.all[0,3] = 'deadlen'
+data.all[,'mean'] = rowMeans(data.all[,4:3603])
+
+#plot(data.all[,'avg'],type='lines')
+
+attach(data.all)
+data.agg = aggregate(mean,list(V1,V2,V3), function(x)mean(x))
+colnames(data.agg) <- c("id","fov","deadlen",'mean')
+
+data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
+data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
+data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
+
+data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
+data.agg[sapply(data.agg['deadlen'],function(x) x==1),'deadlen'] = 'medlen'
+data.agg[sapply(data.agg['deadlen'],function(x) x==2),'deadlen'] = 'upplen'
+
+data.agg[sapply(data.agg['id'],function(x) x==0),'id'] = 'id0'
+#data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
+#data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
+#data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
+#data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
+#data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
+#data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
+
+#data.agg[sapply(data.agg['game'],function(x) x==1),'game'] = '0-1'
+#data.agg[sapply(data.agg['game'],function(x) x==3),'game'] = '2-5'
+#data.agg[sapply(data.agg['game'],function(x) x==10),'game'] = '5-15'
+
+#data.agg[4] = log(data.agg[4])
+#data.agg[17,4] <- -2
+
+detach(data.all)
+
+# one-way repeated measures anova on fov
[…]

### 2009-06-15 trunk/NewResults/timetofailure.py
adding more analysis

@@ -1,0 +1,35 @@
+import sys
+
+done = False
+thresh = 70
+if len(sys.argv) == 2:
+    thresh = int(sys.argv[1])
+
+while not done:
+    for i in xrange(27):
+	line = sys.stdin.readline()
+	if line == "":
+	    done = True
+	    break
+	list = line.rstrip().split(',')
+	id = list[0]
+	fov = list[1]
+	deadlen = list[2]
+	samples = list[3:]
+
+	samples_in = []
+	for j in samples:
+	    try:
+		samples_in.append( float(j) )
+	    except:
+		continue
+
+	#print "%d samples"%len(samples_in)
+
+	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+	ttf = 3600
+	for j in xrange(60*60):
+	    if samples_in[j] > thresh:
+		ttf = j
+		break
+	sys.stdout.write(" %d\n"%ttf)

### 2009-06-15 trunk/NewResults/timetracking.py

@@ -1,0 +1,34 @@
+import sys
+
+done = False
+thresh = 70
+if len(sys.argv) == 2:
+    thresh = int(sys.argv[1])
+
+while not done:
+    for i in xrange(27):
+	line = sys.stdin.readline()
+	if line == "":
+	    done = True
+	    break
+	list = line.rstrip().split(',')
+	id = list[0]
+	fov = list[1]
+	deadlen = list[2]
+	samples = list[3:]
+
+	samples_in = []
+	for j in samples:
+	    try:
+		samples_in.append( float(j) )
+	    except:
+		continue
+
+	#print "%d samples"%len(samples_in)
+
+	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+	totaltime = 0
+	for j in xrange(60*60):
+	    if samples_in[j] <= thresh:
+		totaltime += 1
+	sys.stdout.write(" %d\n"%totaltime)

### 2009-06-15 NewResults/tt.R
adding more analysis

@@ -1,16 +1,16 @@
-datafilename="resampled.csv"
+datafilename="timetracking_20.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 #data.all[0,1] = 
 #data.all[0,2] = 'fov'
 #data.all[0,3] = 'deadlen'
-data.all[,'mean'] = rowMeans(data.all[,4:3603])
+#data.all[,'mean'] = rowMeans(data.all[,4:3603])
 
 #plot(data.all[,'avg'],type='lines')
 
 attach(data.all)
-data.agg = aggregate(mean,list(V1,V2,V3), function(x)mean(x))
-colnames(data.agg) <- c("id","fov","deadlen",'mean')
+data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
+colnames(data.agg) <- c("id","fov","deadlen","tt")
 
 data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
 data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
@@ -30,36 +30,28 @@ data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
 data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id7'
 data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id8'
 
-#data.agg[sapply(data.agg['game'],function(x) x==1),'game'] = '0-1'
-#data.agg[sapply(data.agg['game'],function(x) x==3),'game'] = '2-5'
-#data.agg[sapply(data.agg['game'],function(x) x==10),'game'] = '5-15'
-
-#data.agg[4] = log(data.agg[4])
-#data.agg[17,4] <- -2
-
 detach(data.all)
 
 # one-way repeated measures anova on fov
-#aov.fov=aov(mean~fov+Error(id/fov),data.agg)
+#aov.fov=aov(tt~fov+Error(id/fov),data.agg)
 #print(summary(aov.fov))
-#boxplot(mean~fov,data=data.agg)
+#boxplot(tt~fov,data=data.agg)
 
 
 # one-way repeated measures anova on deadlen
-aov.deadlen = aov(mean ~deadlen+Error(id/deadlen),data.agg)
-print(summary(aov.deadlen))
-boxplot(mean~deadlen,data=data.agg)
[…]

### 2009-06-15 NewResults/ttf.R

@@ -1,16 +1,16 @@
-datafilename="resampled.csv"
+datafilename="timetofailure_20.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 #data.all[0,1] = 
 #data.all[0,2] = 'fov'
 #data.all[0,3] = 'deadlen'
-data.all[,'mean'] = rowMeans(data.all[,4:3603])
+#data.all[,'mean'] = rowMeans(data.all[,4:3603])
 
 #plot(data.all[,'avg'],type='lines')
 
 attach(data.all)
-data.agg = aggregate(mean,list(V1,V2,V3), function(x)mean(x))
-colnames(data.agg) <- c("id","fov","deadlen",'mean')
+data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
+colnames(data.agg) <- c("id","fov","deadlen","ttf")
 
 data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
 data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
@@ -30,13 +30,6 @@ data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
 data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id7'
 data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id8'
 
-#data.agg[sapply(data.agg['game'],function(x) x==1),'game'] = '0-1'
-#data.agg[sapply(data.agg['game'],function(x) x==3),'game'] = '2-5'
-#data.agg[sapply(data.agg['game'],function(x) x==10),'game'] = '5-15'
-
-#data.agg[4] = log(data.agg[4])
-#data.agg[17,4] <- -2
-
 detach(data.all)
 
 # one-way repeated measures anova on fov
@@ -46,20 +39,19 @@ detach(data.all)
 
 
 # one-way repeated measures anova on deadlen
-aov.deadlen = aov(mean ~deadlen+Error(id/deadlen),data.agg)
-print(summary(aov.deadlen))
-boxplot(mean~deadlen,data=data.agg)
+#aov.deadlen = aov(mean ~deadlen+Error(id/deadlen),data.agg)
+#print(summary(aov.deadlen))
+#boxplot(mean~deadlen,data=data.agg)
 
 
 
 # two-way anova repeated measures on both
[…]

### 2009-06-16 NewResults/tt.R
added participant 10

@@ -27,8 +27,9 @@ data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
 data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
 data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
 data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id7'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id8'
+data.agg[sapply(data.agg['id'],function(x) x==7),'id'] = 'id7'
+data.agg[sapply(data.agg['id'],function(x) x==8),'id'] = 'id8'
+data.agg[sapply(data.agg['id'],function(x) x==9),'id'] = 'id9'
 
 detach(data.all)
 
@@ -41,17 +42,17 @@ detach(data.all)
 # one-way repeated measures anova on deadlen
 #aov.deadlen = aov(tt ~deadlen+Error(id/deadlen),data.agg)
 #print(summary(aov.deadlen))
-#boxplot(tt~deadlen,data=data.agg)
+boxplot(tt~deadlen,data=data.agg)
 
 
 
 # two-way anova repeated measures on both
 aov.two = aov(tt ~(fov*deadlen)+Error(id/(fov*deadlen)),data.agg)
 print(summary(aov.two))
-attach(data.agg)
-interaction.plot(fov,deadlen, tt)    #another way to graph the interaction
-detach(data.agg)
-boxplot(tt ~fov*deadlen,data=data.agg)
+#attach(data.agg)
+#interaction.plot(fov,deadlen, tt)    #another way to graph the interaction
+#detach(data.agg)
+#boxplot(tt ~fov*deadlen,data=data.agg)
 
 
 

### 2009-06-17 Paper/paper.tex
starting in on abstract

@@
 \begin{abstract}
+It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.
+We use an “AR simulation” approach, in which a virtual reality (VR) system is used to simulate multiple AR systems.
+In order to validate this approach, we carefully replicated a well-known study by Ellis et al.
+using our simulator, obtaining comparable results.
+
+In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.
+We varied the field of view of the AR display, as well as the reliability of the head tracker.
+In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.
+Our study gives insight into the effect of head tracker reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.
 
@@
 
+motivation??
+
 \section{Related Work}
+
+other studies which vary AR fov, tracking failures
+
+previous X-ray AR work
+
+simulated AR
 

### 2009-06-20 Paper/paper.tex

@@
 
-\title{Title Goes Here}
+\title{A Study of Immersion Factors Using an Augmented Reality Simulation}
 
@@
 
-\section{Experimental Design}
+\section{AR Simulator Validation}
+
+Cha's text goes here
+
+\subsection{Experimental Design}
+
+\subsection{Results}
+
+\subsection{Analysis}
+
+\section{Tracking Study}
 
@@
 
+By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
+Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
+
+\subsection{Experimental Design}
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
 The tracking overlays are  augmented on top of the image to form the simulated augmented reality view.
-
-By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
-Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
 
@@
 
-\section{Results}
+\subsection{Results}
 
-\section{Analysis}
+\subsection{Analysis}
 
-\section{Conclusions}
+\section{Conclusions and Future Work}
 

### 2009-06-21 Paper/paper.tex
working on related works: added description of Livingston and Ai's paper

@@
 
+Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own.
+Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
+A white box was continuously visible, even when the car itself was occluded by a building in the environment.
+Other virtual cars and their associated white boxes acted as distractors.
+At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.
+While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.
+
 other studies which vary AR fov, tracking failures

### 2009-06-22 trunk/NewResults/timetracking.py
made timetracking use factors

@@ -5,6 +5,12 @@ thresh = 70
 if len(sys.argv) == 2:
     thresh = int(sys.argv[1])
 
+id_dict = {}
+for i in xrange(27):
+    id_dict[ "%d"%i ] = "id%d"%i
+fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
+len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+
 while not done:
     for i in xrange(27):
 	line = sys.stdin.readline()
@@ -12,9 +18,9 @@ while not done:
 	    done = True
 	    break
 	list = line.rstrip().split(',')
-	id = list[0]
-	fov = list[1]
-	deadlen = list[2]
+	id = id_dict[list[0].lstrip()]
+	fov = fov_dict[list[1].lstrip()]
+	deadlen = len_dict[list[2].lstrip()]
 	samples = list[3:]
 
 	samples_in = []

### 2009-06-22 NewResults/tt.R
made timetracking use factors

@@ -1,4 +1,5 @@
-datafilename="timetracking_20.csv"
+
+datafilename="timetracking_10_factor.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 #data.all[0,1] = 
@@ -12,73 +13,41 @@ attach(data.all)
 data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
 colnames(data.agg) <- c("id","fov","deadlen","tt")
 
-data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
-data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
-data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
-
-data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
-data.agg[sapply(data.agg['deadlen'],function(x) x==1),'deadlen'] = 'medlen'
-data.agg[sapply(data.agg['deadlen'],function(x) x==2),'deadlen'] = 'upplen'
-
-data.agg[sapply(data.agg['id'],function(x) x==0),'id'] = 'id0'
-data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
-data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
-data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
-data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
-data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
-data.agg[sapply(data.agg['id'],function(x) x==7),'id'] = 'id7'
-data.agg[sapply(data.agg['id'],function(x) x==8),'id'] = 'id8'
-data.agg[sapply(data.agg['id'],function(x) x==9),'id'] = 'id9'
+#data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
+#data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
+#data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
+#
+#data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
[…]

### 2009-06-22 trunk/NewResults/tt.R

@@ -1,4 +1,3 @@
-
 datafilename="timetracking_10_factor.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
@@ -34,20 +33,21 @@ colnames(data.agg) <- c("id","fov","deadlen","tt")
 
 detach(data.all)
 
-data.sorted = data.agg[order(id,deadlen),]
+#data.sorted = data.agg[order(id,deadlen),]
 
 
 
 
 aov.deadlen = aov( tt ~ deadlen + Error(id/deadlen), data.agg )
-aov.fov =     aov( tt ~ fov + Error(id/fov), data.agg )
-aov.two =     aov( tt ~ (fov*deadlen) + Error(id/(fov*deadlen)), data.agg )
+#hsd.deadlen = TukeyHSD( aov.deadlen )
 
-aov2.deadlen = aov( terms( tt ~ deadlen + id, keep.order=T), data.agg )
-aov2.fov =     aov( terms( tt ~ fov, keep.order=T), data.agg )
-aov2.two =     aov( terms( tt ~ fov*deadlen + id, keep.order=T), data.agg )
-
-hsd.deadlen = TukeyHSD( aov.deadlen, "deadlen" )
-hsd.fov = TukeyHSD( aov.fov )
-hsd.two = TukeyHSD( aov.two )
-
+#aov.fov =     aov( tt ~ fov + Error(id/fov), data.agg )
+#aov.two =     aov( tt ~ (fov*deadlen) + Error(id/(fov*deadlen)), data.agg )
+#
+#aov2.deadlen = aov( terms( tt ~ deadlen + id, keep.order=T), data.agg )
+#aov2.fov =     aov( terms( tt ~ fov + id, keep.order=T), data.agg )
+#aov2.two =     aov( terms( tt ~ fov*deadlen + id, keep.order=T), data.agg )
+#
+#hsd.fov = TukeyHSD( aov.fov )
+#hsd.two = TukeyHSD( aov.two )
+#

### 2009-06-22 NewResults/timetracking.py

@@ -18,9 +18,12 @@ while not done:
 	    done = True
 	    break
 	list = line.rstrip().split(',')
-	id = id_dict[list[0].lstrip()]
-	fov = fov_dict[list[1].lstrip()]
-	deadlen = len_dict[list[2].lstrip()]
+	#id = id_dict[list[0].lstrip()]
+	#fov = fov_dict[list[1].lstrip()]
+	#deadlen = len_dict[list[2].lstrip()]
+	id = list[0].lstrip()
+	fov = list[1].lstrip()
+	deadlen = list[2].lstrip()
 	samples = list[3:]
 
 	samples_in = []

### 2009-06-22 trunk/NewResults/mean.R

@@ -1,91 +1,9 @@
-datafilename="resampled.csv"
+datafilename="mean_factor.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
-#data.all[0,1] = 
-#data.all[0,2] = 'fov'
-#data.all[0,3] = 'deadlen'
-data.all[,'mean'] = rowMeans(data.all[,4:3603])
-
-#plot(data.all[,'avg'],type='lines')
+#data.all[,'mean'] = rowMeans(data.all[,4:3603])
 
 attach(data.all)
-data.agg = aggregate(mean,list(V1,V2,V3), function(x)mean(x))
+data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
 colnames(data.agg) <- c("id","fov","deadlen",'mean')
-
-data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
-data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
-data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
-
-data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
-data.agg[sapply(data.agg['deadlen'],function(x) x==1),'deadlen'] = 'medlen'
-data.agg[sapply(data.agg['deadlen'],function(x) x==2),'deadlen'] = 'upplen'
-
-data.agg[sapply(data.agg['id'],function(x) x==0),'id'] = 'id0'
-data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
-data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
-data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
-data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
-data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id7'
-data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id8'
-
-#data.agg[sapply(data.agg['game'],function(x) x==1),'game'] = '0-1'
-#data.agg[sapply(data.agg['game'],function(x) x==3),'game'] = '2-5'
[…]

### 2009-06-22 trunk/NewResults/tt.R

@@ -1,4 +1,4 @@
-datafilename="timetracking_10_factor.csv"
+datafilename="timetracking_5_factor.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 #data.all[0,1] = 
@@ -39,11 +39,11 @@ detach(data.all)
 
 
 aov.deadlen = aov( tt ~ deadlen + Error(id/deadlen), data.agg )
+aov.fov =     aov( tt ~ fov + Error(id/fov), data.agg )
+aov.two =     aov( tt ~ (fov*deadlen) + Error(id/(fov*deadlen)), data.agg )
+
 #hsd.deadlen = TukeyHSD( aov.deadlen )
 
-#aov.fov =     aov( tt ~ fov + Error(id/fov), data.agg )
-#aov.two =     aov( tt ~ (fov*deadlen) + Error(id/(fov*deadlen)), data.agg )
-#
 #aov2.deadlen = aov( terms( tt ~ deadlen + id, keep.order=T), data.agg )
 #aov2.fov =     aov( terms( tt ~ fov + id, keep.order=T), data.agg )
 #aov2.two =     aov( terms( tt ~ fov*deadlen + id, keep.order=T), data.agg )
@@ -51,3 +51,11 @@ aov.deadlen = aov( tt ~ deadlen + Error(id/deadlen), data.agg )
 #hsd.fov = TukeyHSD( aov.fov )
 #hsd.two = TukeyHSD( aov.two )
 #
+
+boxplot( tt ~ deadlen, data.agg )
+boxplot( tt ~ fov, data.agg )
+
+attach(data.agg)
+interaction.plot(fov,deadlen,tt)
+detach(data.agg)
+

### 2009-06-22 NewResults/mean_factor.py

@@ -1,0 +1,36 @@
+import sys
+
+done = False
+
+id_dict = {}
+for i in xrange(27):
+    id_dict[ "%d"%i ] = "id%d"%i
+fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
+len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+
+while not done:
+    for i in xrange(27):
+	line = sys.stdin.readline()
+	if line == "":
+	    done = True
+	    break
+	list = line.rstrip().split(',')
+	id = id_dict[list[0].lstrip()]
+	fov = fov_dict[list[1].lstrip()]
+	deadlen = len_dict[list[2].lstrip()]
+	samples = list[3:]
+
+	samples_in = []
+	for j in samples:
+	    try:
+		samples_in.append( float(j) )
+	    except:
+		continue
+
+	#print "%d samples"%len(samples_in)
+
+	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+	totaltime = 0.
+	for j in xrange(3600):
+	    totaltime += samples_in[j]
+	sys.stdout.write("%lf\n"%(totaltime/3600.))

### 2009-06-22 NewResults/timetracking_factor.py

@@ -1,0 +1,42 @@
+import sys
+
+done = False
+thresh = 70
+if len(sys.argv) == 2:
+    thresh = int(sys.argv[1])
+
+id_dict = {}
+for i in xrange(27):
+    id_dict[ "%d"%i ] = "id%d"%i
+fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
+len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+
+while not done:
+    for i in xrange(27):
+	line = sys.stdin.readline()
+	if line == "":
+	    done = True
+	    break
+	list = line.rstrip().split(',')
+	id = id_dict[list[0].lstrip()]
+	fov = fov_dict[list[1].lstrip()]
+	deadlen = len_dict[list[2].lstrip()]
+	#fov = list[1].lstrip()
+	#deadlen = list[2].lstrip()
+	samples = list[3:]
+
+	samples_in = []
+	for j in samples:
+	    try:
+		samples_in.append( float(j) )
+	    except:
+		continue
+
+	#print "%d samples"%len(samples_in)
+
+	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+	totaltime = 0
+	for j in xrange(60*60):
+	    if samples_in[j] <= thresh:
+		totaltime += 1
+	sys.stdout.write(" %d\n"%totaltime)

### 2009-06-23 trunk/Code/create_extra_conditions.py
adding stuff for extra 3 trials with deadlen = 0.5

@@ -1,0 +1,22 @@
+import pickle
+import random
+from TimelineGen import Timeline
+
+conditions = []
+FOVs = [ 20 ]
+deadLengths = [ 0.5 ]
+
+for fov in FOVs:
+	for deadLength in deadLengths:
+		Timeline.deadLength = deadLength
+		for i in xrange(3):
+			timeline = Timeline()
+			timeline.fov = fov
+			timeline.myDeadLength = deadLength
+			conditions.append(timeline)
+
+random.shuffle(conditions)
+
+file = open("ExtraConditions", 'w')
+pickle.dump(conditions,file)
+file.close()

### 2009-06-23 Code/run_extra.py
adding stuff for extra 3 trials with deadlen = 0.5

@@ -1,0 +1,126 @@
+import viz
+import random
+import viztask
+import math
+import pickle
+import latinSquare
+import environment_setup
+import time
+
+# text box for showing messages
+tbox = viz.addTextbox()	
+tbox.setPosition(0.5,0.35)
+
+# set numtrials number to twenty-seven
+numtrials = 3
+
+# results
+results = []
+
+# load conditions from pickle
+unpicklefile = open('ExtraConditions', 'r')
+conditions = pickle.load(unpicklefile)
+unpicklefile.close()
+
+#for c in conditions:
+#	print c.fov
+#sys.exit()
+
+participantNumber = 0 #change for each participant
+
+resultsPath = "extra_%d_%d.csv"%(participantNumber,time.time())
+
+# function to run the experiment
+def run_tasks():
+	global tbox, message, numtrials, results, conditions, participantNumber, resultsPath
+	#sys.exit()
+	ls = latinSquare.LatinSquare(numtrials)
+	order = ls.getOrder(participantNumber % numtrials)
+	#order[0]=0# temporary debug
+	#print "order:",order
+	#for i in xrange(numtrials):
+	#	print ls.getOrder(i)
+	#sys.exit()
+	
+	#for i in xrange(numtrials):
+	#	print conditions[order[i]].fov
+	#	print conditions[order[i]].myDeadLength
+	
+	# load path set from file
+	unpicklefile = open('pathGen67130', 'r')
+	pg = pickle.load(unpicklefile)
+	unpicklefile.close()
+
+	# show space bar message
+	tbox.message("Press space to start")
+
+	#resultsfile = open(resultsPath,'a')
+	#resultsfile.write("%d\n"%participantNumber)
+	#resultsfile.close()
+
+	# for each trial
+	for i in range(0,numtrials):
+		
+		# wait for space bar
+		yield viztask.waitKeyDown(' ')
+
+		# hide message box
+		tbox.visible(viz.OFF)
+
+		# set up variables for this condition
+		environment_setup.setARfov( conditions[order[i]].fov )
+		
+		# run the path set
+		yield pg.runExperimentPathSet(27+order[i],conditions[order[i]])
+		
+		# append the results
+		# results.append([ncorrect, nfalsepos, nfalseneg])
+		results.append( pg.errlist )
+		print pg.errlist
+		
+		# show trial over message
+		tbox.visible(viz.ON)
+		tbox.message("Trial %d Over"%(i+1))
[…]

### 2009-06-23 Paper/paper.tex
fixed top view image and added it to .tex file.  added a modified room3 model - better for creating figures with (no roof/shorter walls)
began introduction section, added new related works paragraph, added room measurements txt file

@@
 
+Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.
+The ability to see through solid objects has applications in fields ranging from engineering and architecture to military or search-and-rescue operations.
+Unfortunately, implementing x-ray vision is not a trivial task.
+AR systems have cost issules, including those pertaining to level-of-immersion.
+For example, what is the ideal augmented field of view (FOV) for a given scenario?
+Other issues relate to tracking the objects in the x-ray view.
+
 motivation??
@@
 
-Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own.
+Livingston and Ai \cite{4637329} studied the effects of various sources of registration error with an AR simulation similar to our own.
 Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
@@
 While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.
+
+Bane and Hollerer \cite{1383060} propose a set of AR tools for virtual x-ray vision.
+Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
+The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.
 
@@
 \cite{4637329,4811058,1383060}
-
-\section{AR Simulator Validation}
-
-Cha's text goes here
-
-\subsection{Experimental Design}
-
-\subsection{Results}
-
-\subsection{Analysis}
 
@@
 \subsection{Experimental Design}
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=3.5in]{annotated_room.png}
+	\caption{Overhead view of virtual room (approx.
+measure).
+Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.
+Top-hat's initial position is outside the front doorway, directly in the participant's view.}
+\end{figure}
+
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.

### 2009-06-23 trunk/Paper/paperTyler.tex
added some writing

--- 2009-06-21 Paper/paper.tex
@@
 
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=3.5in]{annotated_room.png}
+	\caption{Overhead view of virtual room (approx. measure)}
+\end{figure}
+
 \subsection{Results}
@@
 We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).
+We experimented with longer values but deemed them too long to be realistic through expert analysis.
+This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.
+
+During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.
+The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.
+At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
+The man wearing the tophat had the additional constraints as follows.
+He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.
+In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.
+We made sure these constraints were met by randomly generating paths until they were satisfied.
+We generated 27 sets of patahs this way, one for each trail.
+We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.
+
+The experiment was written in python using the Vizard environemnt.  Tracking was done using IntersenseTracker
+The HMD we used was a Proview HMD with the following attirbutes. ....  Fill this in.
+The computer we used was a Dell with the following specs...........  Fill this in
 

### 2009-06-24 trunk/Paper/paper.tex
outlining
added text about analysis, changed to sections

--- 2009-06-10 Paper/paper.tex
@@
 
+Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.
+The ability to see through solid objects has applications in fields ranging from engineering and architecture to military or search-and-rescue operations.
+Unfortunately, implementing x-ray vision is not a trivial task.
+AR systems have cost issules, including those pertaining to level-of-immersion.
+For example, what is the ideal augmented field of view (FOV) for a given scenario?
+Other issues relate to tracking the objects in the x-ray view.
+
+motivation??
+
 \section{Related Work}
+
+Livingston and Ai \cite{4637329} studied the effects of various sources of registration error with an AR simulation similar to our own.
+Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
+A white box was continuously visible, even when the car itself was occluded by a building in the environment.
+Other virtual cars and their associated white boxes acted as distractors.
+At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.
+While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.
+
+Bane and Hollerer \cite{1383060} propose a set of AR tools for virtual x-ray vision.
+Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
+The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.
+
+other studies which vary AR fov, tracking failures
+
+previous X-ray AR work
+
+simulated AR
 
@@
 
-\section{Experimental Design}
+\section{AR Simulation}
 
@@
 
+By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
+Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
+
+\section{Experimental Design}
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=3.5in]{annotated_room.png}
+	\caption{Overhead view of virtual room (approx.
+measure).
+Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.
+Top-hat's initial position is outside the front doorway, directly in the participant's view.}
+\end{figure}
+
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
 The tracking overlays are  augmented on top of the image to form the simulated augmented reality view.
-
-By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
-Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
 
@@
 We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).
+We experimented with longer values but deemed them too long to be realistic through expert analysis.
+This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.
+
+During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.
+The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.
+At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
+The man wearing the tophat had the additional constraints as follows.
+He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.
+In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.
+We made sure these constraints were met by randomly generating paths until they were satisfied.
+We generated 27 sets of patahs this way, one for each trail.
+We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.
+
+The experiment was written in python using the Vizard environemnt.  Tracking was done using IntersenseTracker
+The HMD we used was a Proview HMD with the following attirbutes. ....  Fill this in.
+The computer we used was a Dell with the following specs...........  Fill this in
+
+\section{Analysis}
+
+In our experiment we define performance as effectiveness in tracking the top hat character.
+We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
+This measure ranges between zero and 180 degrees (we used the smaller of the two options, clockwise and counter-clockwise).
+We record one measurement per video frame, at about sixty frames per second, resulting in about 3600 measurements for a one second trial.
+
+Because the video does not run at exactly 60 Hz, each trial might have a slightly different number of measurements.
+We also record a timestamp for each measurement, so that we can determine their exact frequency.
+As a pre-processing step, we linearly re-sample the data for each trial so that each has exactly 3600 measurements at 60 Hz.
+
+We observed that participants tend to switch between two different states during the experiment: a tracking state, where they are following the target; and a lost state, where they are searching for the target.
+The two states are easily visible in the data.
+In the tracking state, the angular error is generally low.
+We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.
+In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.
+The error may even return to near-zero during a lost state, if the participant unwittingly crosses the target's path.
+
+We hypothesize that less immersive conditions will have longer and more frequent periods where the participant is lost.
+We considered several different metrics which might be appropriate in testing this hypothesis.
+
+The simplest is the average error over an entire trial.
+However, this may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+
+We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.
+When the error rises above the threshold, we consider the participant lost.
+TODO: how to determine this threshold??
+
+One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.
+However a participant may reach the lost state sooner or later depending on the movement of the avatars.
+We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.
+This metric seems to most clearly represent how well a participant performed.
+Trials with either more lost periods or longer lost periods will result in a lower total time tracking.
 
@@
 
-\section{Analysis}
-
-\section{Conclusions}
+\section{Conclusions and Future Work}
 

### 2009-06-24 Paper/paper.tex

@@
 
-\section{Tracking Study}
+\section{AR Simulation}
 
@@
 
-\subsection{Experimental Design}
+\section{Experimental Design}
 
@@
 	\centering
-	\includegraphics[width=3.5in]{annotated_room.png}
+	\includegraphics[width=3.5in]{annotated_roomb.png}
 	\caption{Overhead view of virtual room (approx.
@@
 
-\subsection{Results}
+\section{Analysis}
 
-\subsection{Analysis}
+In our experiment we define performance as effectiveness in tracking the top hat character.
+We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
+This measure ranges between zero and 180 degrees (we used the smaller of the two options, clockwise and counter-clockwise).
+We record one measurement per video frame, at about sixty frames per second, resulting in about 3600 measurements for a one second trial.
+
+Because the video does not run at exactly 60 Hz, each trial might have a slightly different number of measurements.
+We also record a timestamp for each measurement, so that we can determine their exact frequency.
+As a pre-processing step, we linearly re-sample the data for each trial so that each has exactly 3600 measurements at 60 Hz.
+
+We observed that participants tend to switch between two different states during the experiment: a tracking state, where they are following the target; and a lost state, where they are searching for the target.
+The two states are easily visible in the data.
+In the tracking state, the angular error is generally low.
+We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.
+In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.
+The error may even return to near-zero during a lost state, if the participant unwittingly crosses the target's path.
+
+We hypothesize that less immersive conditions will have longer and more frequent periods where the participant is lost.
+We considered several different metrics which might be appropriate in testing this hypothesis.
+
+The simplest is the average error over an entire trial.
+However, this may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+
+We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.
+When the error rises above the threshold, we consider the participant lost.
+TODO: how to determine this threshold??
+
+One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.
+However a participant may reach the lost state sooner or later depending on the movement of the avatars.
+We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.
+This metric seems to most clearly represent how well a participant performed.
+Trials with either more lost periods or longer lost periods will result in a lower total time tracking.
+
+\section{Results}
 

### 2009-06-24 Paper/paper.tex

@@
 Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.
-The ability to see through solid objects has applications in fields ranging from engineering and architecture to military or search-and-rescue operations.
+The ability to see through solid objects has applications in various fields ranging from engineering and architecture to military or search-and-rescue operations.
 Unfortunately, implementing x-ray vision is not a trivial task.
@@
 For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view.
+Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors, where tracking is temporarily lost.
+
+In this paper, we study the effects of varying augmented FOV along with object-tracking dropouts.
+By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.
+The simulation is a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
+We wish to answer the following questions.
+What effect does varying the augmented FOV have on user tracking performance?
+What effect do different dropout lengths have on performance?
+What are the interactions between these two variables?
 

### 2009-06-24 trunk/Paper/paper.tex
filled in some technical details

@@
 \begin{abstract}
-It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.
+%It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.
 We use an “AR simulation” approach, in which a virtual reality (VR) system is used to simulate multiple AR systems.
@@
 For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view.
+Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.
 
-motivation??
+In this paper, we study the effects of varying augmented FOV along with object-tracking dropouts.
+By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.
+The simulation is a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
+We wish to answer the following questions.
+What effect does varying the augmented FOV have on user tracking performance?
+What effect do different dropout lengths have on performance?
+What are the interactions between these two variables?
 
@@
 	\centering
-	\includegraphics[width=3.5in]{annotated_room.png}
+	\includegraphics[width=3.5in]{figures/augmentedroom.png}
 	\caption{Overhead view of virtual room (approx.
@@
 
-The experiment was written in python using the Vizard environemnt.  Tracking was done using IntersenseTracker
-The HMD we used was a Proview HMD with the following attirbutes. ....  Fill this in.
-The computer we used was a Dell with the following specs...........  Fill this in
+The experiment was written in Python using the Vizard environment.
+Tracking was done using InterSense InertiaCube3 orientation tracker which has 180 Hz update rate.
+We used a Pro-View 60 head mounted display which displays $640\times480$ video at 60 Hz, and has 36 degrees vertical field of view.
+We did not use stereoscopic display.
+We ran the experiment on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP.
 

### 2009-06-24 trunk/NewResults/timetracking_factor.py
new way of timetracking so you dont say its lost until for a number of seconds

@@ -1,7 +1,10 @@
 import sys
 
 done = False
-thresh = 70
+
+thresh = 12
+secondsLoss = 2
+
 if len(sys.argv) == 2:
     thresh = int(sys.argv[1])
 
@@ -11,32 +14,61 @@ for i in xrange(27):
 fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
 len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
 
+resamples = open("resampled.csv")
+
 while not done:
     for i in xrange(27):
-	line = sys.stdin.readline()
-	if line == "":
-	    done = True
-	    break
-	list = line.rstrip().split(',')
-	id = id_dict[list[0].lstrip()]
-	fov = fov_dict[list[1].lstrip()]
-	deadlen = len_dict[list[2].lstrip()]
-	#fov = list[1].lstrip()
-	#deadlen = list[2].lstrip()
-	samples = list[3:]
+		
+		line = resamples.readline()
+		if line == "":
+			done = True
+			break
+		list = line.rstrip().split(',')
+		id = id_dict[list[0].lstrip()]
+		fov = fov_dict[list[1].lstrip()]
+		deadlen = len_dict[list[2].lstrip()]
+		#fov = list[1].lstrip()
+		#deadlen = list[2].lstrip()
+		samples = list[3:]
 
-	samples_in = []
-	for j in samples:
-	    try:
-		samples_in.append( float(j) )
-	    except:
-		continue
+		samples_in = []
+		for j in samples:
+			try:
+				samples_in.append( float(j) )
+			except:
+				continue
 
-	#print "%d samples"%len(samples_in)
+		#print "%d samples"%len(samples_in)
 
-	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
-	totaltime = 0
-	for j in xrange(60*60):
-	    if samples_in[j] <= thresh:
-		totaltime += 1
-	sys.stdout.write(" %d\n"%totaltime)
+		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+		totaltime = 0
+		
+		numLoss = 0
+		reachedThreshold = 0
+		totaltime = 60*60
+		
+		for j in xrange(60*60):
+			if samples_in[j] > thresh:
+				numLoss += 1
+				
+				if numLoss >= (secondsLoss*60):
+					if reachedThreshold == 0:
[…]

### 2009-06-24 trunk/Paper/paper.tex

@@
 \begin{abstract}
-%It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.
-We use an “AR simulation” approach, in which a virtual reality (VR) system is used to simulate multiple AR systems.
-In order to validate this approach, we carefully replicated a well-known study by Ellis et al.
-using our simulator, obtaining comparable results.
-
+This is a study using a virtual reality (VR) simulation system to examine the effects of various problems associated with augmented reality (AR).
 In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.
@@
 
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=3.5in]{figures/tophatscreenshot.png}
+	\caption{Sample user-view of AR simulation.
+Top-hat is visible (non-occluded) in the doorway.
+The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
+\end{figure}
+
 \section{Results}

### 2009-06-24 Paper/paper.tex
created figures directory

@@
 \begin{abstract}
-It is extremely challenging to run controlled studies comparing multiple augmented reality (AR) systems.
-We use an “AR simulation” approach, in which a virtual reality (VR) system is used to simulate multiple AR systems.
-In order to validate this approach, we carefully replicated a well-known study by Ellis et al.
-using our simulator, obtaining comparable results.
-
+This is a study using a virtual reality (VR) simulation system to examine the effects of various problems associated with augmented reality (AR).
 In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.
@@
 For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors, where tracking is temporarily lost.
+Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.
 
@@
 What are the interactions between these two variables?
-
-motivation??
 
@@
 	\centering
-	\includegraphics[width=3.5in]{annotated_roomb.png}
+	\includegraphics[width=3.5in]{figures/augmentedroom.png}
 	\caption{Overhead view of virtual room (approx.
@@
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
-The tracking overlays are  augmented on top of the image to form the simulated augmented reality view.
+The tracking overlays are augmented on top of the original image to form the simulated augmented reality view.
 
@@
 In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.
-Each trial lasted sixty seconds, and had seven tracking failures.
-The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).
+Each trial is 60 seconds, with seven tracking failures.
+The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
+The length of tracking failures varied between two seconds (longest), one second (medium failure) and zero seconds (no failure).
 We experimented with longer values but deemed them too long to be realistic through expert analysis.
@@
 
-During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.
-The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.
-At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
-The man wearing the tophat had the additional constraints as follows.
+During each trial there are 20 people in addition to a virtual man (Tophat) wearing a tall black tophat walking an unpredictable path outside of the room.
+The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
+Each path had the constraint that they must stay within the rectangular area of size X by X surrounding the room.
+At each path point, the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
+\emph{Tophat} had the additional constraints as follows.
 He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.
@@
 
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=3.5in]{figures/tophatscreenshot.png}
+	\caption{Sample user-view of AR simulation.
+\emph{Tophat} is visible (non-occluded) in the doorway.
+The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
+\end{figure}
+
 \section{Results}

### 2009-06-24 trunk/NewResults/resampler_extra.py
changed resampler.py so for to work with the extra data

@@ -1,0 +1,82 @@
+import sys
+
+class Resampler:
+    def __init__(self,samples,timing,rate):
+		self.timestep = 1. / rate
+		self.samples = samples
+		self.timing = timing
+		self.samples.insert(0,samples[0])
+		self.timing.insert(0,0)
+		lastsample = self.samples[len(self.samples)-1]
+		lasttiming = self.timing[len(self.timing)-1]
+		for i in xrange(100):
+			self.samples.append(lastsample)
+			lasttiming += self.timestep
+			self.timing.append(lasttiming)
+		self.time = 0
+    
+    def step(self):
+		for i in xrange( len(self.samples)-1 ):
+			if self.time >= self.timing[i] and self.time <= self.timing[i+1]:
+				t0 = self.timing[i]
+				t1 = self.timing[i+1]
+				s0 = self.samples[i]
+				s1 = self.samples[i+1]
+				alpha = ( self.time - t0 ) / ( t1 - t0 )
+				break
+		sample = alpha * s0 + ( 1. - alpha ) * s1
+		self.time += self.timestep
+		return (self.time,sample)
+
+for i in xrange(27):
+    samples_line = sys.stdin.readline()
+    timing_line = sys.stdin.readline()
+    list = samples_line.rstrip().split(',')
+    id = int(list[0]) + 12
+    fov = list[1]
+    deadlen = list[2]
+    samples = list[3:]
+    list = timing_line.rstrip().split(',')
+    timing = list[3:]
+
+    samples_in = []
+    timing_in = []
+    for j in samples:
+		try:
+			samples_in.append( float(j) )
+		except:
+			continue
+    for j in timing:
+		try:
+			timing_in.append( float(j) )
+		except:
+			continue
+
+	if int(fov) == 20:
+
+		#print "%d samples, %d timings"%(len(samples_in),len(timing_in))
+		rs = Resampler( samples_in, timing_in, 60. )
+
+		samples_out = []
+		timing_out = []
+
+		
+		sys.stdout.write("%s,%s,%s,"%(str(id),fov,deadlen))
+		for j in xrange(60*60):
+			(time,sample) = rs.step()
+			sys.stdout.write("%f"%sample)
+			if j != 3599:
+				sys.stdout.write(",")
+			#samples_out.append(sample)
+			#timing_out.append(time)
+		sys.stdout.write("\n")
+    
+
+
+
+
+
+
+
+
+    

### 2009-06-24 NewResults/timetracking_delay_factor.py
change for added delay

@@ -1,0 +1,74 @@
+import sys
+
+done = False
+
+thresh = 12
+secondsLoss = 2
+
+if len(sys.argv) == 2:
+    thresh = int(sys.argv[1])
+
+id_dict = {}
+for i in xrange(27):
+    id_dict[ "%d"%i ] = "id%d"%i
+fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
+len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+
+resamples = open("resampled.csv")
+
+while not done:
+    for i in xrange(27):
+		
+		line = resamples.readline()
+		if line == "":
+			done = True
+			break
+		list = line.rstrip().split(',')
+		id = id_dict[list[0].lstrip()]
+		fov = fov_dict[list[1].lstrip()]
+		deadlen = len_dict[list[2].lstrip()]
+		#fov = list[1].lstrip()
+		#deadlen = list[2].lstrip()
+		samples = list[3:]
+
+		samples_in = []
+		for j in samples:
+			try:
+				samples_in.append( float(j) )
+			except:
+				continue
+
+		#print "%d samples"%len(samples_in)
+
+		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+		totaltime = 0
+		
+		numLoss = 0
+		reachedThreshold = 0
+		totaltime = 60*60
+		
+		for j in xrange(60*60):
+			if samples_in[j] > thresh:
+				numLoss += 1
+				
+				if numLoss >= (secondsLoss*60):
+					if reachedThreshold == 0:
+						totaltime -= numLoss
+						reachedThreshold = 1
+					else:
+						totaltime -= 1
+			else:
+				numLoss = 0
+				reachedThreshold = 0
+		sys.stdout.write(" %d\n"%totaltime)
+		
+		
+		
+		
+		
+		
+		
+		
+		
+		
+		

### 2009-06-24 NewResults/timetracking_factor.py
return to version without delay

@@ -1,10 +1,7 @@
 import sys
 
 done = False
-
 thresh = 12
-secondsLoss = 2
-
 if len(sys.argv) == 2:
     thresh = int(sys.argv[1])
 
@@ -18,57 +15,30 @@ resamples = open("resampled.csv")
 
 while not done:
     for i in xrange(27):
-		
-		line = resamples.readline()
-		if line == "":
-			done = True
-			break
-		list = line.rstrip().split(',')
-		id = id_dict[list[0].lstrip()]
-		fov = fov_dict[list[1].lstrip()]
-		deadlen = len_dict[list[2].lstrip()]
-		#fov = list[1].lstrip()
-		#deadlen = list[2].lstrip()
-		samples = list[3:]
+	line =  resamples.readline()
+	if line == "":
+	    done = True
+	    break
+	list = line.rstrip().split(',')
+	id = id_dict[list[0].lstrip()]
+	fov = fov_dict[list[1].lstrip()]
+	deadlen = len_dict[list[2].lstrip()]
+	#fov = list[1].lstrip()
+	#deadlen = list[2].lstrip()
+	samples = list[3:]
 
-		samples_in = []
-		for j in samples:
-			try:
-				samples_in.append( float(j) )
-			except:
-				continue
+	samples_in = []
+	for j in samples:
+	    try:
+		samples_in.append( float(j) )
+	    except:
+		continue
 
-		#print "%d samples"%len(samples_in)
+	#print "%d samples"%len(samples_in)
 
-		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
-		totaltime = 0
-		
-		numLoss = 0
-		reachedThreshold = 0
-		totaltime = 60*60
-		
-		for j in xrange(60*60):
-			if samples_in[j] > thresh:
-				numLoss += 1
-				
-				if numLoss >= (secondsLoss*60):
-					if reachedThreshold == 0:
-						totaltime -= numLoss
-						reachedThreshold = 1
-					else:
-						totaltime -= 1
-			else:
-				numLoss = 0
-				reachedThreshold = 0
-		sys.stdout.write(" %d\n"%totaltime)
-		
-		
-		
-		
-		
-		
-		
-		
-		
-		
-		
+	sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+	totaltime = 0
+	for j in xrange(60*60):
+	    if samples_in[j] <= thresh:
+		totaltime += 1
+	sys.stdout.write(" %d\n"%totaltime)

### 2009-06-24 NewResults/resampler_extra.py

@@ -32,7 +32,7 @@ for i in xrange(27):
     samples_line = sys.stdin.readline()
     timing_line = sys.stdin.readline()
     list = samples_line.rstrip().split(',')
-    id = int(list[0]) + 12
+    id = int(list[0]) - 12
     fov = list[1]
     deadlen = list[2]
     samples = list[3:]
@@ -52,14 +52,13 @@ for i in xrange(27):
 		except:
 			continue
 
-	if int(fov) == 20:
+    if int(fov) == 20:
 
 		#print "%d samples, %d timings"%(len(samples_in),len(timing_in))
 		rs = Resampler( samples_in, timing_in, 60. )
 
 		samples_out = []
 		timing_out = []
-
 		
 		sys.stdout.write("%s,%s,%s,"%(str(id),fov,deadlen))
 		for j in xrange(60*60):

### 2009-06-24 NewResults/numtimeslost_factor.py
for checking the number of times abe was lost

@@ -1,0 +1,72 @@
+import sys
+
+done = False
+
+thresh = 12
+secondsLoss = 2
+
+if len(sys.argv) == 2:
+    thresh = int(sys.argv[1])
+
+id_dict = {}
+for i in xrange(27):
+    id_dict[ "%d"%i ] = "id%d"%i
+fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
+len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+
+resamples = open("resampled.csv")
+
+while not done:
+    for i in xrange(27):
+		
+		line = resamples.readline()
+		if line == "":
+			done = True
+			break
+		list = line.rstrip().split(',')
+		id = id_dict[list[0].lstrip()]
+		fov = fov_dict[list[1].lstrip()]
+		deadlen = len_dict[list[2].lstrip()]
+		#fov = list[1].lstrip()
+		#deadlen = list[2].lstrip()
+		samples = list[3:]
+
+		samples_in = []
+		for j in samples:
+			try:
+				samples_in.append( float(j) )
+			except:
+				continue
+
+		#print "%d samples"%len(samples_in)
+
+		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
+		totaltime = 0
+		
+		numLoss = 0
+		reachedThreshold = 0
+		timeslost = 0
+		
+		for j in xrange(60*60):
+			if samples_in[j] > thresh:
+				numLoss += 1
+				
+				if numLoss >= (secondsLoss*60):
+					if reachedThreshold == 0:
+						timeslost += 1
+						reachedThreshold = 1
+			else:
+				numLoss = 0
+				reachedThreshold = 0
+		sys.stdout.write(" %d\n"%timeslost)
+		
+		
+		
+		
+		
+		
+		
+		
+		
+		
+		

### 2009-06-24 trunk/Paper/paper.tex

@@
 \author{
-%Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
-%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
-%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research}
+Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
+\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
+\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research
 }
@@
 
+The design of the experiment is based around the idea that a task should have straightforward quantitative results.
+Also, the task should be reasonably simple so as not to frustrate the participants.
+Added challenges were that the task should be generic so as to be generizable, yet grounded in reality (i.e.
+not an abstract world), to enhance participant familiarity with the AR task.
+
+We developed a tracking scenario, where the user is asked to visually track a virtual person as he moves around the scene.
+The participant is centered inside a room with various doors and windows giving a view to the outside world.
+They may change their orientation as the please, but not change position (3 degrees of freedom).
+The user's view to the tracked virtual person may be occluded at any time by the walls of the building.
+The augmented view element is composed of red marker rectangles indicating the location of the tracked person and other virtual people.
+
 \begin{figure}[ht!]
 	\centering
-	\includegraphics[width=3.5in]{figures/augmentedroom.png}
+	\includegraphics[width=3in]{figures/augmentedroom.png}
 	\caption{Overhead view of virtual room (approx.
@@
 In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.
-Each trial lasted sixty seconds, and had seven tracking failures.
-The total vertical field of view of our HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-We varied the length of tracking failures between two seconds, one second and zero seconds (no failure).
+Each trial is 60 seconds, with seven tracking failures.
+The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
+The length of tracking failures varied between two seconds (high failure), one second (medium failure) and zero seconds (no failure).
 We experimented with longer values but deemed them too long to be realistic through expert analysis.
@@
 
-During each trial there are 20 people in addtition to the man wearing the tophat waking around outside of the room.
-The walking path for each person was given by a randomly generated set of points with the constraint that they must stay within the rectangular area of size X by X surrounding the room.
-At each point in the path the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
-The man wearing the tophat had the additional constraints as follows.
-He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.
-In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.
-We made sure these constraints were met by randomly generating paths until they were satisfied.
-We generated 27 sets of patahs this way, one for each trail.
-We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.
+During each trial there are 20 people in addition to a virtual man \emph{Tophat} wearing a tall black tophat walking an unpredictable path outside of the room.
+The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
+Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
+At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
+\emph{Tophat} had the additional constraints as follows.
+He must start at the same location for each trial which made him visible by the participant through the front door the building.
+In order to make the paths by of similar difficulty \emph{Tophat} must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
+These constraints were met by randomly generating paths until all were satisfied.
+We generated 27 sets of paths in this fashion, each corresponding to a specific trial.
+We used a latin squares ordering to discourage the effects of learning.
 
-The experiment was written in Python using the Vizard environment.
-Tracking was done using InterSense InertiaCube3 orientation tracker which has 180 Hz update rate.
-We used a Pro-View 60 head mounted display which displays $640\times480$ video at 60 Hz, and has 36 degrees vertical field of view.
-We did not use stereoscopic display.
-We ran the experiment on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP.
+We implemented this experiment with Python and the Vizard virtual reality toolkit.
+Head-tracking was handled by a InterSense InertiaCube3 orientation tracker with a 180 Hz refresh rate.
+We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
+As we were not concerned with stereo performance, we turned off the stereoscopic feature.
+The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
 
@@
 
-In our experiment we define performance as effectiveness in tracking the top hat character.
+In our experiment we define performance as effectiveness in tracking \emph{Tophat}.
 We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
@@
 The simplest is the average error over an entire trial.
-However, this may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+This may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
 
@@
 When the error rises above the threshold, we consider the participant lost.
-TODO: how to determine this threshold??
+Here, we use a 12 degree error as the threshold.
 
@@
 	\centering
-	\includegraphics[width=3.5in]{figures/tophatscreenshot.png}
+	\includegraphics[width=3in]{figures/tophatscreenshot.png}
 	\caption{Sample user-view of AR simulation.
-Top-hat is visible (non-occluded) in the doorway.
+\emph{Tophat} is visible (non-occluded) in the doorway.
 The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}

### 2009-06-24 trunk/Paper/paper.tex

@@
 \begin{abstract}
-This is a study using a virtual reality (VR) simulation system to examine the effects of various problems associated with augmented reality (AR).
-In a second study using the AR simulator, we studied a tracking task where the AR interface provides x-ray vision to help track a moving target.
+
+Augmented reality (AR) systems are expensive.
+It is worth studying the effects of level of immersion in AR systems in order to determine an appropriate cost to benefit ration.
+This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion in AR.
+We analyze a tracking task where the AR interface provides x-ray vision to help track a moving target.
 We varied the field of view of the AR display, as well as the reliability of the head tracker.
@@
 \section*{Acknowledgements}
+Cha Lee, Steffan Gauglitz, funding, and all of our participants.
 

### 2009-06-24 Paper/paper.tex
some comments about analysis

@@
 The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-The length of tracking failures varied between two seconds (longest), one second (medium failure) and zero seconds (no failure).
+The length of tracking failures varied between two seconds (high failure), one second (medium failure) and zero seconds (no failure).
 We experimented with longer values but deemed them too long to be realistic through expert analysis.
@@
 
-During each trial there are 20 people in addition to a virtual man (Tophat) wearing a tall black tophat walking an unpredictable path outside of the room.
+During each trial there are 20 people in addition to a virtual man \emph{Tophat} wearing a tall black tophat walking an unpredictable path outside of the room.
 The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
-Each path had the constraint that they must stay within the rectangular area of size X by X surrounding the room.
-At each path point, the person would randomly change his walking speed to a new value which was within a range we deemed accecptable through expert analysis.
+Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
+At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
 \emph{Tophat} had the additional constraints as follows.
-He must start at the same location for each trial which made him visible by the participant through the door shaped opening in the building.
-In order to make the paths by of similar dificulty we decided, through expert analysis, that through the trial he must have walked around the entire building at least once, and he must be visible through the windows for at least $x$ seconds.
-We made sure these constraints were met by randomly generating paths until they were satisfied.
-We generated 27 sets of patahs this way, one for each trail.
-We used a balanced latin square to order these paths and the conditions to discourage the effects of learning.
+He must start at the same location for each trial which made him visible by the participant through the front door the building.
+In order to make the paths by of similar difficulty \emph{Tophat} must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
+These constraints were met by randomly generating paths until all were satisfied.
+We generated 27 sets of paths in this fashion, each corresponding to a specific trial.
+We used a latin squares ordering to discourage the effects of learning.
 
-The experiment was written in Python using the Vizard environment.
-Tracking was done using InterSense InertiaCube3 orientation tracker which has 180 Hz update rate.
-We used a Pro-View 60 head mounted display which displays $640\times480$ video at 60 Hz, and has 36 degrees vertical field of view.
-We did not use stereoscopic display.
-We ran the experiment on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP.
+We implemented this experiment with Python and the Vizard virtual reality toolkit.
+Head-tracking was handled by a InterSense InertiaCube3 orientation tracker with a 180 Hz refresh rate.
+We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
+As we were not concerned with stereo performance, we turned off the stereoscopic feature.
+The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
 
@@
 The simplest is the average error over an entire trial.
-However, this may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+This may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+
+% median ??
 
@@
 
+% count number of times lost
+
+% time tracking is overview metric -- captures both number of times lost and length of lost periods
+% number of times lost is piece of overview -- number of times lost.
+
 \begin{figure}[ht!]
 	\centering
-	\includegraphics[width=3.5in]{figures/tophatscreenshot.png}
+	\includegraphics[width=3in]{figures/tophatscreenshot.png}
 	\caption{Sample user-view of AR simulation.
-\emph{Tophat} is visible (non-occluded) in the doorway.
+\emph{Top Hat} is visible (non-occluded) in the doorway.
 The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
@@
 
+% should we include data from older generation??
+
+% look at analysis from mean.
+
+% time tracking: fov, deadlen significant.  interaction is weird....
+
+% num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+
 \section{Conclusions and Future Work}

### 2009-06-24 trunk/Paper/paper.tex
working on analysis section

@@
 
-Livingston and Ai \cite{4637329} studied the effects of various sources of registration error with an AR simulation similar to our own.
+Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own \cite{4637329}.
 Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
@@
 
-Bane and Hollerer \cite{1383060} propose a set of AR tools for virtual x-ray vision.
+Bane and H\"{o}llerer propose a set of AR tools for virtual x-ray vision \cite{1383060}.
 Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
@@
 
-other studies which vary AR fov, tracking failures
+% TODO: add discussion of validity of AR simulation w/ citations
 
-previous X-ray AR work
-
-simulated AR
-
-\cite{4637329,4811058,1383060}
+%other studies which vary AR fov, tracking failures?
 
@@
 
-\begin{figure}[ht!]
+\begin{figure}[t]
 	\centering
@@
 
-During each trial there are 20 people in addition to a virtual man \emph{Tophat} wearing a tall black tophat walking an unpredictable path outside of the room.
+During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat walking an unpredictable path outside of the room.
 The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
@@
 At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
-\emph{Tophat} had the additional constraints as follows.
+\emph{Top Hat} had the additional constraints as follows.
 He must start at the same location for each trial which made him visible by the participant through the front door the building.
@@
 The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3in]{figures/tophatscreenshot.png}
+	\caption{Sample user-view of AR simulation.
+\emph{Top Hat} is visible (non-occluded) in the doorway.
+The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
+\end{figure}
 
@@
 
-We hypothesize that less immersive conditions will have longer and more frequent periods where the participant is lost.
+% TODO: include two plots of error here (one with losses, one without)
+
+We hypothesize that less immersive conditions will have longer and more frequent periods where the participant loses the target.
 We considered several different metrics which might be appropriate in testing this hypothesis.
@@
 
-We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.
-When the error rises above the threshold, we consider the participant lost.
-Here, we use a 12 degree error as the threshold.
+% median ??
 
-One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.
-However a participant may reach the lost state sooner or later depending on the movement of the avatars.
-We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.
-This metric seems to most clearly represent how well a participant performed.
+We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.
+The threshold specifies the maximum angular error that still represents successful tracking of the target.
+%When the error rises above the threshold, we consider the participant lost.
+We use a threshold of one quarter of the total horizontal field of view of the HMD, $\tau = 12$ degrees.
+% TODO: how to motivate this threshold?
+
+Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.
+However, this may not be a very descriptive measure, since a participant may reach the lost state sooner or later depending on the movement of the avatars, which varies between trials.
+
+We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.
+This metric seems to most generally represent how well a participant performed.
 Trials with either more lost periods or longer lost periods will result in a lower total time tracking.
 
-\begin{figure}[ht!]
+For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial.
+We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 0.5 seconds.
+% TODO: how to motivate this threshold?
+
+%time tracking is overview metric -- captures both number of times lost and length of lost periods
+
+%number of times lost is piece of overview -- number of times lost.
+
+\section{Results}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_deadlen.pdf}
+%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods.}
+%\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_fov.pdf}
+%	\caption{Box plot of the number of times the participant becomes lost versus the field of view of the augmentations.}
+%\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
+%	\caption{Interaction plot of the effect of field of view and tracking failure on the number of times the participant becomes lost.}
+%\end{figure}
+
+\begin{figure}[t]
 	\centering
-	\includegraphics[width=3in]{figures/tophatscreenshot.png}
-	\caption{Sample user-view of AR simulation.
-\emph{Tophat} is visible (non-occluded) in the doorway.
-The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
+	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
+	\caption{\label{fig:tt_deadlen}Box plot of the amount of time that the target is near the crosshairs versus the length of tracker failure periods.}
 \end{figure}
 
-\section{Results}
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/tt_fov.pdf}
+%	\caption{Box plot of time the participant is not lost versus the field of view of the augmentations.}
+%\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
+%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
+%\end{figure}
+
+% TODO: should we include data from older generation??
+
+% TODO: look at analysis from mean.
+
+We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.
+Table \ref{fig:anova} gives the results of this analysis.
+Both the failure length and the field of view had a very significant effect on performance using this metric.
+% TODO: interaction?
+
+% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+
+Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus failure length.
+Performance decreases between zero and one seconds, but seems to level off between one and two.
+We ran a post-hoc analysis to determine which conditions were significantly different.
+Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).
+This suggests that we may be seeing a thresholding effect above one second tracking failures.
+
+This suggests that we have reached a threshold at one second.
+We had seven participants complete an extra three trials with the medium field of view and tracking failure periods of 0.5 seconds.
+Figure 8 shows the data.
+There wasn't a significant difference between the zero length and half second length conditions.
+For future work, we could find the threshold between 0.5 and 1.
+
+%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
+
+\begin{figure}[htb]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
+Metric & Factor & F value & Pr \\
+\hline
+Time Tracking & Failure Length & 20.41 & $<0.001$\\
+Time Tracking & Field of View & 13.549 & $<0.001$\\
+Number of Times Lost & Failure Length & 61.816 & $<0.001$\\
+\hline
+\end{tabular}
+ \caption{\label{fig:anova}Factors which have a significant effect.}
+\end{figure}
 

### 2009-06-24 Paper/paper.tex

@@
 Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.
-The ability to see through solid objects has applications in various fields ranging from engineering and architecture to military or search-and-rescue operations.
+The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality}, medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.
 Unfortunately, implementing x-ray vision is not a trivial task.
@@
 
-We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.
-When the error rises above the threshold, we consider the participant lost.
-TODO: how to determine this threshold??
+We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.
+The threshold specifies the maximum angular error that still represents successful tracking of the target.
+%When the error rises above the threshold, we consider the participant lost.
+We use a threshold of one quarter of the total horizontal field of view of the HMD, $\tau = 12$ degrees.
+% TODO: how to motivate this threshold?
 
-One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.
-However a participant may reach the lost state sooner or later depending on the movement of the avatars.
-We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.
-This metric seems to most clearly represent how well a participant performed.
+Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.
+However, this may not be a very descriptive measure, since a participant may reach the lost state sooner or later depending on the movement of the avatars, which varies between trials.
+
+We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.
+This metric seems to most generally represent how well a participant performed.
 Trials with either more lost periods or longer lost periods will result in a lower total time tracking.
 
-% count number of times lost
+For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial.
+We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 0.5 seconds.
+% TODO: how to motivate this threshold?
 
-% time tracking is overview metric -- captures both number of times lost and length of lost periods
-% number of times lost is piece of overview -- number of times lost.
+%time tracking is overview metric -- captures both number of times lost and length of lost periods
 
-\begin{figure}[ht!]
-	\centering
-	\includegraphics[width=3in]{figures/tophatscreenshot.png}
-	\caption{Sample user-view of AR simulation.
-\emph{Top Hat} is visible (non-occluded) in the doorway.
-The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
-\end{figure}
+%number of times lost is piece of overview -- number of times lost.
 
@@
 
-% should we include data from older generation??
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_deadlen.pdf}
+%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods.}
+%\end{figure}
 
-% look at analysis from mean.
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_fov.pdf}
+%	\caption{Box plot of the number of times the participant becomes lost versus the field of view of the augmentations.}
+%\end{figure}
 
-% time tracking: fov, deadlen significant.  interaction is weird....
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
+%	\caption{Interaction plot of the effect of field of view and tracking failure on the number of times the participant becomes lost.}
+%\end{figure}
 
-% num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
+	\caption{\label{fig:tt_deadlen}Box plot of the amount of time that the target is near the crosshairs versus the length of tracker failure periods.}
+\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/tt_fov.pdf}
+%	\caption{Box plot of time the participant is not lost versus the field of view of the augmentations.}
+%\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
+%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
+%\end{figure}
+
+% TODO: should we include data from older generation??
+
+% TODO: look at analysis from mean.
+
+We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.
+Table \ref{fig:anova} gives the results of this analysis.
+Both the failure length and the field of view had a very significant effect on performance using this metric.
+% TODO: interaction?
+
+% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+
+Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus failure length.
+Performance decreases between zero and one seconds, but seems to level off between one and two.
+We ran a post-hoc analysis to determine which conditions were significantly different.
+Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).
+This suggests that we may be seeing a thresholding effect above one second tracking failures.
+
+This suggests that we have reached a threshold at one second.
+We had seven participants complete an extra three trials with the medium field of view and tracking failure periods of 0.5 seconds.
+Figure 8 shows the data.
+There wasn't a significant difference between the zero length and half second length conditions.
+For future work, we could find the threshold between 0.5 and 1.
+
+%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
+
+\begin{figure}[htb]
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
+Metric & Factor & F value & Pr \\
+\hline
+Time Tracking & Failure Length & 20.41 & $<0.001$\\
+Time Tracking & Field of View & 13.549 & $<0.001$\\
+Number of Times Lost & Failure Length & 61.816 & $<0.001$\\
+\hline
+\end{tabular}
+ \caption{\label{fig:anova}Factors which have a significant effect.}
+\end{figure}
 

### 2009-06-24 trunk/Paper/paper.tex

@@
 
-In this paper, we study the effects of varying augmented FOV along with object-tracking dropouts.
+In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.
+Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
+It also gives us the ability to remove unnecessary factors from generating noise in our data: illumination
+
 By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.
@@
 
-Bane and H\"{o}llerer propose a set of AR tools for virtual x-ray vision \cite{1383060}.
+Bane and Hollerer propose a set of AR tools for virtual x-ray vision \cite{1383060}.
 Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
@@
 
-The design of the experiment is based around the idea that a task should have straightforward quantitative results.
+The design of the experiment is built with several goals in mind.
+The task should have straightforward, quantitative results.
 Also, the task should be reasonably simple so as not to frustrate the participants.
-Added challenges were that the task should be generic so as to be generizable, yet grounded in reality (i.e.
+Further, the task should be generic so as to be generizable, yet grounded in reality (i.e.
 not an abstract world), to enhance participant familiarity with the AR task.
 
-We developed a tracking scenario, where the user is asked to visually track a virtual person as he moves around the scene.
+We developed a tracking task scenario, where the user is asked to visually track a virtual person as it moves throughout the scene.
 The participant is centered inside a room with various doors and windows giving a view to the outside world.
 They may change their orientation as the please, but not change position (3 degrees of freedom).
-The user's view to the tracked virtual person may be occluded at any time by the walls of the building.
+The virtual person to be tracked walks an unpredictable path outside the room.
+The user's view to the virtual person may be occluded at any time by the walls of the building; at other times the virtual person may be visible through doors or windows in the building.
 The augmented view element is composed of red marker rectangles indicating the location of the tracked person and other virtual people.
@@
 
-Because the video does not run at exactly 60 Hz, each trial might have a slightly different number of measurements.
+Because the video does not always run at precisely 60 Hz, each trial might have a slightly different number of measurements.
 We also record a timestamp for each measurement, so that we can determine their exact frequency.
@@
 The simplest is the average error over an entire trial.
-This may not accurately capture the amount that a participant is lost, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+This may not accurately represent tracking failures, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
 
@@
 This metric seems to most generally represent how well a participant performed.
-Trials with either more lost periods or longer lost periods will result in a lower total time tracking.
+Trials with either more or longer lost periods will result in a lower total time tracking.
 
@@
 %	\includegraphics[width=3in]{figures/numtimes_deadlen.pdf}
-%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods.}
+%	\caption{Box plot of the number of times the participant is lost versus the length of the sensor dropouts.}
 %\end{figure}
@@
 %	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
-%	\caption{Interaction plot of the effect of field of view and tracking failure on the number of times the participant becomes lost.}
+%	\caption{Interaction plot of the effect of field of view and registration dropouts on the number of times the participant becomes lost.}
 %\end{figure}
@@
 	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
-	\caption{\label{fig:tt_deadlen}Box plot of the amount of time that the target is near the crosshairs versus the length of tracker failure periods.}
+	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
 \end{figure}
@@
 
-Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus failure length.
+Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
 Performance decreases between zero and one seconds, but seems to level off between one and two.
@@
 Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).
-This suggests that we may be seeing a thresholding effect above one second tracking failures.
+This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
 
 This suggests that we have reached a threshold at one second.
-We had seven participants complete an extra three trials with the medium field of view and tracking failure periods of 0.5 seconds.
+We had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
 Figure 8 shows the data.
@@
 \hline
-Time Tracking & Failure Length & 20.41 & $<0.001$\\
+Time Tracking & Dropout Length & 20.41 & $<0.001$\\
 Time Tracking & Field of View & 13.549 & $<0.001$\\
-Number of Times Lost & Failure Length & 61.816 & $<0.001$\\
+Number of Times Lost & Dropout Length & 61.816 & $<0.001$\\
 \hline

### 2009-06-24 Paper/paper.tex

@@
 
-In this paper, we study the effects of varying augmented FOV along with object-tracking dropouts.
-By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.
-The simulation is a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
+In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.
+Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
+It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
+
+The AR simulation is built as a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
 We wish to answer the following questions.

### 2009-06-24 trunk/Paper/paper.tex
more figures

@@
 Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
-It also gives us the ability to remove unnecessary factors from generating noise in our data: illumination
+It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
 
-By using an AR simulation built with virtual reality, it is possible to simulate the tracking dropouts, which would otherwise be difficult or impossible to replicate.
-The simulation is a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
+The AR simulation is built as a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
 We wish to answer the following questions.
@@
 
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
-%	\caption{Interaction plot of the effect of field of view and registration dropouts on the number of times the participant becomes lost.}
-%\end{figure}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3.5in]{figures/tt_deadlen.pdf}
+	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
+\end{figure}
+% TODO: make the units seconds and change the labels on the bottom of this figure
 
@@
 	\centering
-	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
-	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
+	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
+	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
 \end{figure}
@@
 
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
-%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
-%\end{figure}
-
 % TODO: should we include data from older generation??
@@
 We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.
-Table \ref{fig:anova} gives the results of this analysis.
+Figure \ref{fig:anova} gives the results of this analysis.
 Both the failure length and the field of view had a very significant effect on performance using this metric.
@@
 
+We also used a two-way {ANOVA} to examine the number of times the participant loses the target.
 % TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
-
-Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
-Performance decreases between zero and one seconds, but seems to level off between one and two.
-We ran a post-hoc analysis to determine which conditions were significantly different.
-Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).
-This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
-
-This suggests that we have reached a threshold at one second.
-We had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
-Figure 8 shows the data.
-There wasn't a significant difference between the zero length and half second length conditions.
-For future work, we could find the threshold between 0.5 and 1.
+\begin{figure}[h]
 
@@
 
-\begin{figure}[htb]
 \centering
@@
 \hline
-Time Tracking & Dropout Length & 20.41 & $<0.001$\\
-Time Tracking & Field of View & 13.549 & $<0.001$\\
-Number of Times Lost & Dropout Length & 61.816 & $<0.001$\\
+Time Tracking & Failure Length & 28.904 & $<0.001$\\
+Time Tracking & Field of View & 15.800 & $<0.001$\\
+Number of Times Lost & Failure Length & 8.5008 & $<0.001$\\
+Number of Times Lost & Field of View & 13.060 & $<0.001$\\
 \hline
@@
 
+Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
+Performance decreases between zero and one seconds, but seems to level off between one and two.
+We ran a post-hoc analysis to determine which conditions were significantly different.
+Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p=??), but not significantly different from each other (p=??).
+This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
+% TODO: table with values
+
+To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
+Figure \ref{fig:extra} shows a box plot of time tracking versus sensor dropout length for these seven participants.
+Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.
+%TODO: table with values
+With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
+%For future work, we could find the threshold between 0.5 and 1.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3.5in]{figures/extra.pdf}
+	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected this data from seven participants, using the medium field of view.}
+\end{figure}
+
+%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
+
 \section{Conclusions and Future Work}
@@
 \section*{Acknowledgements}
-Cha Lee, Steffan Gauglitz, funding, and all of our participants.
+Cha Lee, Steffen Gauglitz, funding, and all of our participants.
 

### 2009-06-24 Paper/paper.tex

@@
 
-Augmented reality (AR) systems are expensive.
-It is worth studying the effects of level of immersion in AR systems in order to determine an appropriate cost to benefit ration.
-This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion in AR.
-We analyze a tracking task where the AR interface provides x-ray vision to help track a moving target.
+This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion and tracking sensor reliability in AR.
+We analyze participant performance on a tracking task in which the AR interface provides x-ray vision to help track a moving target.
 We varied the field of view of the AR display, as well as the reliability of the head tracker.
 In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.
-Our study gives insight into the effect of head tracker reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.
+Our study gives insight into the effect of tracking sensor reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.
 
@@
 A participant stands in the middle of a square room with doors and windows on all sides.
-Outside of the room, people are walking around the building.
-The target person to be tracked is visually distinguishable by a large black top hat.
+Virtual people walk unpredictable paths outside the building.
+The target to be tracked is one of these people, visually distinguishable by a large black top hat.
 The augmented reality interface overlays a translucent red rectangle on each person, which is visible even when the person is occluded by a wall.
-However, the target person's overlay is not different from any other person's overlay; the target is only identifiable when visible through a window or a door.
+However, the target person's overlay is identical to the other virtual people overlays; the target may only be distinguished when visible through a window or a door.
 
-To actually run this experiment using a true augmented reality system would be difficult.
+To actually run this experiment using a true augmented reality system would be extremely difficult.
 Many confederates would be needed to walk around the room, and their movements would need to be accurately reproduced for each new participant.
 Also, the confederates would need to be continously tracked for the augmented overlays to be displayed.
-Instead, we simulated the experiment setup in a completely virtual environment.
+Instead, we chose to simulate the experiment setup in a completely virtual environment.
 
 By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
-Using a simulated environment also makes the experiment more controllable, and thus more easily reproduced.
+This simulated environment also affords easily controlled variables, and thus it may easily be reproduced.
 
@@
 
-The parameters of our experiment are divided between the ``real'' immersion factors, such as the field of view of the HMD, and the ``augmented'' immersion factors, such as the field of view of the simulated AR display, and the performance of the head tracker.
-We introduce periods of tracking failure, where the augmented overlays disappear, to simulate the effect of failures in a real tracking system.
-For example, such failures might occur with a magnetic tracker near interfering materials, or with visual tracking when tracked features are lost.
+The parameters of our experiment are divided between the ``real'' immersion factors, such as the field of view of the HMD, and the ``augmented'' immersion factors, such as the field of view of the simulated AR display, and the performance of the tracking sensor.
+We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of failures in a real tracking system.
+For example, such dropouts might occur with a magnetic tracker near interfering materials, or with visual tracking when tracked features are lost.
 
@@
 The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-The length of tracking failures varied between two seconds (high failure), one second (medium failure) and zero seconds (no failure).
-We experimented with longer values but deemed them too long to be realistic through expert analysis.
+The length of dropouts varied between two seconds (highest), one second (medium) and zero seconds (lowest).
+We experimented with longer dropout periods but deemed them too long.
 This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.
@@
 We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
-As we were not concerned with stereo performance, we turned off the stereoscopic feature.
+As we were not concerned the effects of stereo, we turned off the stereoscopic feature.
 The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
@@
 % TODO: how to motivate this threshold?
-
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3.5in]{figures/tt_deadlen.pdf}
+	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
+\end{figure}
 Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.
@@
 
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
-%	\caption{Interaction plot of the effect of field of view and registration dropouts on the number of times the participant becomes lost.}
-%\end{figure}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
-	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
-\end{figure}
+% TODO: make the units seconds and change the labels on the bottom of this figure
 
@@
 
+We also used a two-way {ANOVA} to examine the number of times the participant loses the target.
 % TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
+	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
+\end{figure}
 
-Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
-Performance decreases between zero and one seconds, but seems to level off between one and two.
-We ran a post-hoc analysis to determine which conditions were significantly different.
-Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).
-This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
-
-This suggests that we have reached a threshold at one second.
-We had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
-Figure 8 shows the data.
-There wasn't a significant difference between the zero length and half second length conditions.
-For future work, we could find the threshold between 0.5 and 1.
-
-%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
-
-\begin{figure}[htb]
+\begin{figure}[h]
 \centering
@@
 
+Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
+Performance decreases between zero and one seconds, but seems to level off between one and two.
+We ran a post-hoc analysis to determine which conditions were significantly different.
+Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p=??), but not significantly different from each other (p=??).
+This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
+% TODO: table with values
+
+To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
+Figure \ref{fig:extra} shows a box plot of time tracking versus sensor dropout length for these seven participants.
+Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.
+%TODO: table with values
+With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
+%For future work, we could find the threshold between 0.5 and 1.
+
+\begin{figure}[htb]
+	\centering
+	\includegraphics[width=3.5in]{figures/extra.pdf}
+	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected this data from seven participants, using the medium field of view.}
+\end{figure}
+
+%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
+
 \section{Conclusions and Future Work}

### 2009-06-25 trunk/Paper/paper.tex
laying out figures
Added information about participants

@@
 We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
-As we were not concerned with stereo performance, we turned off the stereoscopic feature.
+As we were not concerned the effects of stereo, we turned off the stereoscopic feature.
 The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
+
+We ran this experiment on 19 subjects, 11 were male, and 8 female.
+The participants ranged in age from 23 to 59, 15 were in the range 23 to 30, and the remaining 4 in the range 51 to 49.
+None were colorblind.
+Most users were heavy computer users with at least somewhat familiarity with virtual reality and augmented reality.
+Only one participant played 3d video games often.
+After every 9 trials we had the participants take a mandatory 5 minute break.
+Some experienced light fatigue, but none experienced any strong effects of motion sickness.
 
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=3.5in]{figures/tt_deadlen.pdf}
-	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
-\end{figure}
 % TODO: make the units seconds and change the labels on the bottom of this figure
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
-	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
-\end{figure}
 
@@
 % TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
+	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
+\end{figure}
+
 \begin{figure}[h]
-
-%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.
-
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
 	\centering

### 2009-06-25 trunk/Paper/paper.tex

@@
 By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
-This simulated environment also affords easily controlled variables, and thus it may easily be reproduced.
+This simulated environment also affords easily controlled variables, and thus the experiment may be replicated.
 
@@
 At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
-\emph{Top Hat} had the additional constraints as follows.
+Top Hat had the additional constraints as follows.
 He must start at the same location for each trial which made him visible by the participant through the front door the building.
-In order to make the paths by of similar difficulty \emph{Tophat} must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
+In order to make the paths by of similar difficulty Top Hat must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
 These constraints were met by randomly generating paths until all were satisfied.
@@
 
-In our experiment we define performance as effectiveness in tracking \emph{Tophat}.
+In our experiment we define performance as effectiveness in tracking Top Hat.
 We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.

### 2009-06-25 Paper/paper.tex
Added subsection for participants

@@
 The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
+
+\subsection(Participants)
+
+We ran this experiment on 19 subjects, 11 were male, and 8 female.
+The participants ranged in age from 23 to 59, 15 were in the range 23 to 30, and the remaining 4 in the range 51 to 49.
+None were colorblind.
+Most users were heavy computer users with at least somewhat familiarity with virtual reality and augmented reality.
+Only one participant played 3D video games often.
+After every 9 trials we had the participants take a mandatory 5 minute break.
+Some experienced light fatigue, but none experienced any strong effects of motion sickness.
 

### 2009-06-25 trunk/Paper/paper.tex
adding more tex, made it anonymous

@@
 
-\title{A Study of Immersion Factors Using an Augmented Reality Simulation}
+\title{A Study of an X-Ray Vision Tracking Task in Simulated Augmented Reality}
 
@@
 \author{
-Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
-\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research
+Submission 191
+%Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
+%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
+%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research
 }
@@
 
+\subsection{Implementation}
+
 We implemented this experiment with Python and the Vizard virtual reality toolkit.
@@
 
+\subsection{Participants}
+
 We ran this experiment on 19 subjects, 11 were male, and 8 female.
@@
 Most users were heavy computer users with at least somewhat familiarity with virtual reality and augmented reality.
-Only one participant played 3d video games often.
+Only one participant played 3D video games often.
 After every 9 trials we had the participants take a mandatory 5 minute break.
@@
 Figure \ref{fig:anova} gives the results of this analysis.
-Both the failure length and the field of view had a very significant effect on performance using this metric.
+Both the failure length and the field of view have a very significant effect on performance using this metric.
 % TODO: interaction?
 
-We also used a two-way {ANOVA} to examine the number of times the participant loses the target.
-% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.
+We also used a two-way {ANOVA} to examine the number of times the participant loses the target, with results shown in Figure \ref{fig:anova}.
+Both field of view and dropout length have a significant effect.
+There also is an interaction between the two factors.
+Figure \ref{fig:interaction} shows a plot of this interaction.
+
+From this plot we can see that with a small field of view, performance is equally bad with no dropouts or one second dropouts.
+The small field of view seems to have a dominating effect on performance.
+When the field of view is increased to the medium level, performance dramatically increases when there are no dropouts, but only gets slightly better when there are dropouts.
+Finally, we do not observe much difference in performance between the medium and large field of view cases.
+Using Tukey's HSD to compare means, the medium and large field of view conditions are in fact not significantly different (p=??).
+
 \begin{figure}[t]
@@
 	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
-	\caption{Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
+	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
 \end{figure}
@@
 
-\section*{Acknowledgements}
-Cha Lee, Steffen Gauglitz, funding, and all of our participants.
+%\section*{Acknowledgements}
+%Cha Lee, Steffen Gauglitz, funding, and all of our participants.
 

### 2009-06-25 Paper/paper.tex

@@
 This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion and tracking sensor reliability in AR.
-We analyze participant performance on a tracking task in which the AR interface provides x-ray vision to help track a moving target.
-We varied the field of view of the AR display, as well as the reliability of the head tracker.
+We analyze participant performance on a target following task in which the AR interface provides x-ray vision to help track a moving target.
+We varied the field of view of the AR display, as well as the reliability of the head tracking sensor.
 In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.
-Our study gives insight into the effect of tracking sensor reliability on user performance in a tracking task, as well as the relationship between reliability and field of view in an AR system.
+Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between reliability and field of view in an AR system.
 
@@
 
-In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.
+In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a target following task.
 Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
@@
 
-The AR simulation is built as a tracking task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
+The AR simulation is built as a target following task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
 We wish to answer the following questions.
-What effect does varying the augmented FOV have on user tracking performance?
+What effect does varying the augmented FOV have on user task performance?
 What effect do different dropout lengths have on performance?
@@
 
-Our experiment involves a tracking task which uses an augmented reality interface to allow people to be tracked behind walls.
+Our experiment involves a target following task which uses an augmented reality interface to allow people to be tracked behind walls.
 A participant stands in the middle of a square room with doors and windows on all sides.
@@
 
-We developed a tracking task scenario, where the user is asked to visually track a virtual person as it moves throughout the scene.
+We developed a target following task scenario, where the user is asked to visually track a virtual person as it moves throughout the scene.
 The participant is centered inside a room with various doors and windows giving a view to the outside world.
@@
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
-The tracking overlays are augmented on top of the original image to form the simulated augmented reality view.
+The augmented overlays are merged with the original image to form the simulated augmented reality view.
 
 The parameters of our experiment are divided between the ``real'' immersion factors, such as the field of view of the HMD, and the ``augmented'' immersion factors, such as the field of view of the simulated AR display, and the performance of the tracking sensor.
-We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of failures in a real tracking system.
-For example, such dropouts might occur with a magnetic tracker near interfering materials, or with visual tracking when tracked features are lost.
+We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.
+For example, such dropouts might occur with a magnetic sensor near interfering material.
 
-In our experiment we vary two parameters: the field of view of the AR interface and the length of tracking failure periods.
-Each trial is 60 seconds, with seven tracking failures.
+In our experiment we vary two parameters: the field of view of the AR interface and the length of sensor dropout periods.
+Each trial is 60 seconds, with seven sensor dropouts.
 The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
@@
 
-\subsection(Participants)
+\subsection{Participants}
 
@@
 	\caption{Sample user-view of AR simulation.
-\emph{Top Hat} is visible (non-occluded) in the doorway.
+Top Hat is visible (non-occluded) in the doorway.
 The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
@@
 
-In our experiment we define performance as effectiveness in tracking Top Hat.
+In our experiment we define performance as effectiveness in following Top Hat.
 We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
 This measure ranges between zero and 180 degrees (we used the smaller of the two options, clockwise and counter-clockwise).
-We record one measurement per video frame, at about sixty frames per second, resulting in about 3600 measurements for a one second trial.
+We record one measurement per video frame, at about sixty frames per second, resulting in approximately 3600 measurements for a one minute trial.
 
@@
 
-We observed that participants tend to switch between two different states during the experiment: a tracking state, where they are following the target; and a lost state, where they are searching for the target.
+We observed that participants tend to switch between two different states during the experiment: a target following state, and a lost state, where they are searching for the target.
 The two states are easily visible in the data.
-In the tracking state, the angular error is generally low.
+In the target following state, the angular error is generally low.
 We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.
@@
 The simplest is the average error over an entire trial.
-This may not accurately represent tracking failures, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the tracking state.
+This may not accurately represent sensor dropouts, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the target following state.
 
@@
 We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.
-The threshold specifies the maximum angular error that still represents successful tracking of the target.
+The threshold specifies the maximum angular error that still represents successful following of the target.
 %When the error rises above the threshold, we consider the participant lost.
@@
 
-We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.
+We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the target following state.
 This metric seems to most generally represent how well a participant performed.
-Trials with either more or longer lost periods will result in a lower total time tracking.
+Trials with either more or longer lost periods will result in a lower total time following the target.
 
@@
 
-Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus dropout length.
+Figure \ref{fig:tt_deadlen} shows a box plot of time following target versus dropout length.
 Performance decreases between zero and one seconds, but seems to level off between one and two.
@@
 To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
-Figure \ref{fig:extra} shows a box plot of time tracking versus sensor dropout length for these seven participants.
+Figure \ref{fig:extra} shows a box plot of time following target versus sensor dropout length for these seven participants.
 Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.

### 2009-06-25 Paper/paper.tex

@@
 For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.
+Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where position information is temporarily lost.
 
@@
 Virtual people walk unpredictable paths outside the building.
-The target to be tracked is one of these people, visually distinguishable by a large black top hat.
+The target to be followed is one of these people, visually distinguishable by a large black top hat.
 The augmented reality interface overlays a translucent red rectangle on each person, which is visible even when the person is occluded by a wall.
@@
 
-By simulating augmented reality, we are freed from the current restrictions of tracking and display technology.
+By simulating augmented reality, we are freed from the current restrictions of display technology.
 This simulated environment also affords easily controlled variables, and thus the experiment may be replicated.
@@
 
-We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the target following state.
+We also consider a measurement we call ``Time Following Target,'' which is the amount of time spent in the target following state.
 This metric seems to most generally represent how well a participant performed.
@@
 \hline
-Time Tracking & Failure Length & 28.904 & $<0.001$\\
-Time Tracking & Field of View & 15.800 & $<0.001$\\
-Number of Times Lost & Failure Length & 8.5008 & $<0.001$\\
+Time Following Target & Dropout Length & 28.904 & $<0.001$\\
+Time Following Target & Field of View & 15.800 & $<0.001$\\
+Number of Times Lost & Dropout Length & 8.5008 & $<0.001$\\
 Number of Times Lost & Field of View & 13.060 & $<0.001$\\

### 2009-06-25 trunk/Paper/paper.tex
Added extra intro parpagraph

@@
 
-Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.
+Augmented Reality systems may be used to simulate x-ray vision.
 The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality}, medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.
@@
 For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.
+Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where position information is temporarily lost.
 
-In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a tracking task.
+A task common to many AR x-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
+Problems with the AR system tracking objects accurately is often an issue which is likely to have an effect on task performance.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the augmented field of view s well as its interaction with tracking difficulties.
+We created a task that aimed to accomplish these goals in a generic fashion in order to be generalizable while still being grounded in reality.
+
+In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a target following task.
 Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
@@
 
-Our experiment involves a tracking task which uses an augmented reality interface to allow people to be tracked behind walls.
+Our experiment involves a target following task which uses an augmented reality interface to allow people to be tracked behind walls.
 A participant stands in the middle of a square room with doors and windows on all sides.
 Virtual people walk unpredictable paths outside the building.
-The target to be tracked is one of these people, visually distinguishable by a large black top hat.
+The target to be followed is one of these people, visually distinguishable by a large black top hat.
 The augmented reality interface overlays a translucent red rectangle on each person, which is visible even when the person is occluded by a wall.
@@
 
-We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.
+We also consider a measurement we call ``Time Following Target,'' which is the amount of time spent in the target following state.
 This metric seems to most generally represent how well a participant performed.
-Trials with either more or longer lost periods will result in a lower total time tracking.
+Trials with either more or longer lost periods will result in a lower total time following the target.
 
@@
 \bibliographystyle{acmsiggraph}
-\nocite{*}
 \bibliography{paper}

### 2009-06-25 Paper/paper.tex

@@
 
-\title{A Study of an X-Ray Vision Tracking Task in Simulated Augmented Reality}
+\title{A Study of Target Following with X-Ray Vision in 
+%Simulated
+Augmented Reality}
 
@@
 \author{
-Submission 191
+Submission 165
 %Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
@@
 Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where position information is temporarily lost.
+
+A task common to many AR x-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
+Problems with the AR system tracking objects accurately is often an issue which is likely to have an effect on task performance.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the augmented field of view s well as its interaction with tracking difficulties.
+We created a task that aimed to accomplish these goals in a generic fashion in order to be generalizable while still being grounded in reality.
 
@@
 	\centering
-	\includegraphics[width=3in]{figures/tophatscreenshot.png}
+	\includegraphics[width=3in]{figures/view_comp.png}
 	\caption{Sample user-view of AR simulation.

### 2009-06-25 Paper/paper.tex
multiple view figure caption fixed

@@
 	\includegraphics[width=3in]{figures/view_comp.png}
-	\caption{Sample user-view of AR simulation.
-Top Hat is visible (non-occluded) in the doorway.
-The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
+	\caption{Sample views of AR simulation.
+\emph{Upper-left:} Detail of Top Hat in foreground (posed view).
+\emph{Upper-right:} A user view of Top Hat (non-occluded) in the doorway.
+Red rectangles overlay the people, indicating their location.
+\emph{Lower-left:} Three distractor people are visible through a large window.
+\emph{Lower-right:} A typical user view of a corner occluding several people.}
 \end{figure}

### 2009-06-25 Paper/paper.tex

--- 2009-06-25 Paper/paper.tex
@@
 
-\begin{figure}[t]
+\begin{figure*}[ht!]
 	\centering
-	\includegraphics[width=3in]{figures/view_comp.png}
-	\caption{Sample user-view of AR simulation.
-Top Hat is visible (non-occluded) in the doorway.
-The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
-\end{figure}
+	\includegraphics[width=7in]{figures/triple_view.png}
+	\caption{Sample views of AR simulation.
+Transparent red rectangles overlay the people, indicating their location.
+\emph{Left:} A user view of Top Hat (non-occluded) in the doorway.
+\emph{Middle:} Three distractor people are visible through a large window.
+\emph{Right:} A typical user view of a corner occluding several people.}
+\end{figure*}
 

### 2009-06-25 Paper/paper.tex

@@
 
-The parameters of our experiment are divided between the ``real'' immersion factors, such as the field of view of the HMD, and the ``augmented'' immersion factors, such as the field of view of the simulated AR display, and the performance of the tracking sensor.
+The parameters of our experiment are divided between the ``real'' components of immersion, such as the field of view of the HMD, and the ``augmented'' components of immersion, such as the field of view of the simulated AR display, and the performance of the tracking sensor.
 We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.

### 2009-06-25 Paper/paper.tex
Added descriptions of why we chose the thresholds

@@
 
-\title{A Study of Target Following with X-Ray Vision in 
-%Simulated
-Augmented Reality}
+\title{A Study of an X-Ray Vision Tracking Task in Simulated Augmented Reality}
 
@@
 \author{
-Submission 165
+Submission 191
 %Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
@@
 %When the error rises above the threshold, we consider the participant lost.
-We use a threshold of one quarter of the total horizontal field of view of the HMD, $\tau = 12$ degrees.
+While conducting the trials we noticed that the participants spent a majority of the time correctly following Top Hat, computing the median angular error at each time-step across all trials gave us values mostly around 4 degrees so we consider values close to this as successful tracking.
+We also noticed that during trials as Top Hat would change direction the participant's angular error would increase a small amount for a short time, without losing the target.
+As a result we doubled the 4 degrees, giving us a threshold of $\tau = 8$ degrees
 % TODO: how to motivate this threshold?
@@
 \end{figure}
-Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.
+Using this threshold we measure a participant's \emph{time to failure}, the time until the a lost state is encountered.
 However, this may not be a very descriptive measure, since a participant may reach the lost state sooner or later depending on the movement of the avatars, which varies between trials.
 
-We also consider a measurement we call ``Time Following Target,'' which is the amount of time spent in the target following state.
+We also consider a measurement we call \emph{time following target}, which is the amount of time spent successfully following the target.
 This metric seems to most generally represent how well a participant performed.
@@
 
-For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial.
-We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 0.5 seconds.
+For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial, or  \emph{number of times lost}.
+We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 4 seconds.
+A 4 second threshold was used because it is double the maximum dropout length, allowing the user a few seconds to resume successful tracking after the dropout before they are considered to have lost the target.
 % TODO: how to motivate this threshold?

### 2009-06-25 trunk/Paper/paper.tex
changing title

@@
 This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion and tracking sensor reliability in AR.
-We analyze participant performance on a target following task in which the AR interface provides x-ray vision to help track a moving target.
+We analyze participant performance on a target following task in which the AR interface provides X-ray vision to help track a moving target.
 We varied the field of view of the AR display, as well as the reliability of the head tracking sensor.
@@
 
-Augmented Reality systems may be used to simulate x-ray vision.
+Augmented Reality systems may be used to simulate X-ray vision.
 The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality}, medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.
-Unfortunately, implementing x-ray vision is not a trivial task.
-AR systems have cost issules, including those pertaining to level-of-immersion.
-For example, what is the ideal augmented field of view (FOV) for a given scenario?
-Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where position information is temporarily lost.
+Unfortunately, implementing X-ray vision is not a trivial task.
+AR systems certain unsolved issues, including those pertaining to tracking the objects in the X-ray view.
+For example, GPS dropout errors \cite{4079263} may cause augmented position information to be temporarily lost.
+We hypothesize that sensor dropouts will decrease user performance, but this problem may be mitigated by increasing the augmented field of view (FOV).
 
-A task common to many AR x-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
+A task common to many AR X-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
 Problems with the AR system tracking objects accurately is often an issue which is likely to have an effect on task performance.
@@
 
-Bane and Hollerer propose a set of AR tools for virtual x-ray vision \cite{1383060}.
+Bane and Hollerer propose a set of AR tools for virtual X-ray vision \cite{1383060}.
 Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
@@
 
-\begin{figure}[t]
+\begin{figure*}[ht!]
 	\centering
-	\includegraphics[width=3in]{figures/tophatscreenshot.png}
-	\caption{Sample user-view of AR simulation.
-Top Hat is visible (non-occluded) in the doorway.
-The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
-\end{figure}
+	\includegraphics[width=7in]{figures/triple_view.png}
+	\caption{Sample views of AR simulation.
+Transparent red rectangles overlay the people, indicating their location.
+\emph{Left:} A user view of Top Hat (non-occluded) in the doorway.
+\emph{Middle:} Three distractor people are visible through a large window.
+\emph{Right:} A typical user view of a corner occluding several people.}
+\end{figure*}
 

### 2009-06-25 Paper/paper.tex
new title!

@@
 
-\title{A Study of an X-Ray Vision Tracking Task in Simulated Augmented Reality}
+\title{Evaluating the Effects of Tracker Reliability and Field of View on a Target Folling Task in Augmented Reality}
 

### 2009-06-25 Paper/paper.tex

@@
 
-\title{Evaluating the Effects of Tracker Reliability and Field of View on a Target Folling Task in Augmented Reality}
+\title{Evaluating the Effects of Tracker Reliability and Field of View on a Target Following Task in Augmented Reality}
 
@@
 
-This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion and tracking sensor reliability in AR.
-We analyze participant performance on a target following task in which the AR interface provides X-ray vision to help track a moving target.
-We varied the field of view of the AR display, as well as the reliability of the head tracking sensor.
-In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.
+We examine the effects of level of immersion and tracking sensor reliability in augmented reality (AR) X-ray vision with a virtual reality (VR) based simulation.
+We analyze participant performance on a target following task while varying the field of view of the AR display, as well as the reliability of the head tracking sensor.
+In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
 Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between reliability and field of view in an AR system.
@@
 
+Ragan et.
+al.
+\cite{4811058} developed an AR simulation using virtual reality to examine the effect of registration errors on an object manipulation tasks.
+The implemented task involved moving a virtual ring from one end of an irregularly shaped tube to the other, while avoiding collisions between the two objects.
+By using VR to simulate this task, they were able to independently control tracking jitter and latency variables in a pilot study.
+
 % TODO: add discussion of validity of AR simulation w/ citations
@@
 \section{AR Simulation}
+
+\begin{figure*}[ht!]
+	\centering
+	\includegraphics[width=7in]{figures/triple_view.png}
+	\caption{Sample views of AR simulation.
+Transparent red rectangles overlay the people, indicating their location.
+\emph{Left:} A user view of Top Hat (non-occluded) in the doorway.
+\emph{Middle:} Three distractor people are visible through a large window.
+\emph{Right:} A typical user view of a corner occluding several people.}
+\end{figure*}
 
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=3in]{figures/augmentedroom.png}
-	\caption{Overhead view of virtual room (approx.
-measure).
-Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.
-Top-hat's initial position is outside the front doorway, directly in the participant's view.}
-\end{figure}
-
 Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=3in]{figures/augmentedroom.png}
+	\caption{Overhead view of virtual room (approx.
+measure).
+Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.
+Top-hat's initial position is outside the front doorway, directly in the participant's view.}
+\end{figure}
+
 \subsection{Implementation}
@@
 Some experienced light fatigue, but none experienced any strong effects of motion sickness.
-
-\begin{figure*}[ht!]
-	\centering
-	\includegraphics[width=7in]{figures/triple_view.png}
-	\caption{Sample views of AR simulation.
-Transparent red rectangles overlay the people, indicating their location.
-\emph{Left:} A user view of Top Hat (non-occluded) in the doorway.
-\emph{Middle:} Three distractor people are visible through a large window.
-\emph{Right:} A typical user view of a corner occluding several people.}
-\end{figure*}
 

### 2009-06-25 trunk/Paper/paper.tex

@@
 
-\title{A Study of an X-Ray Vision Tracking Task in Simulated Augmented Reality}
+\title{Evaluating the Effects of Tracker Reliability and Field of View on a Target Following Task in Augmented Reality}
 
@@
 
-This is an AR study, simulated using a virtual reality (VR), to examine the effects of level of immersion and tracking sensor reliability in AR.
-We analyze participant performance on a target following task in which the AR interface provides X-ray vision to help track a moving target.
-We varied the field of view of the AR display, as well as the reliability of the head tracking sensor.
-In low reliability conditions, we simulate tracker failure by disabling the augmented view of the scene for brief periods.
-Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between reliability and field of view in an AR system.
+We examine the effects of level of immersion and tracking sensor reliability in augmented reality (AR) X-ray vision with a virtual reality (VR) based simulation.
+We analyze participant performance on a target following task while varying the field of view of the AR display, as well as the reliability of the head tracking sensor.
+In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
+Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between sensor reliability and field of view in an AR system.
 

### 2009-06-25 trunk/Paper/paper.tex
added conclusions

@@
 	\centering
-	\includegraphics[width=3.5in]{figures/tt_deadlen.pdf}
-	\caption{\label{fig:tt_deadlen}Box plot of the amount of time the target is near the crosshairs versus the length of sensor dropouts.}
+	\includegraphics[width=1.5in]{figures/tt_deadlen.pdf}
+	\includegraphics[width=1.5in]{figures/tt_fov.pdf}\\
+	\includegraphics[width=1.5in]{figures/numtimes_deadlen.pdf}
+	\includegraphics[width=1.5in]{figures/numtimes_fov.pdf}
+	\caption{\label{fig:tt_deadlen}Box plots for our two independent variables and two metrics.
+In general, performance increases when the length of sensor dropout periods is lessened, and when the AR field of view increases.}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=3.5in]{figures/numtimes_interaction.pdf}
+	\includegraphics[width=2.5in]{figures/numtimes_interaction.pdf}
 	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
@@
 	\centering
-	\includegraphics[width=3.5in]{figures/extra.pdf}
+	\includegraphics[width=2.5in]{figures/extra.pdf}
 	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.
@@
 
+We designed an experiment to study a target following task which makes use of X-ray vision through an augmented reality device.
+We used an augmented reality simulation to make the experiment feasible.
+
+We found that for our task, a low field of view has an adverse effect on performance.
+This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.
+With reliable head-based rendering, a medium field of view is enough for good performance.
+We did not see a significant performance increase when increasing to a high field of view.
+
+Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.
+However, for our task, the performance decrease levels off at one second dropouts.
+Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.
+This suggests that below a certain length of dropouts, good level of performance can be maintained, provided that the field of view is large enough.
+This threshold on the dropouts would change depending on what is being followed, but we expect the same pattern would appear in other target following scenarios.
+
+Future work lies in further investigating the effect of the dropout length 
+below the one second level.  Also, this work points to other interesting experiments which might test the interaction between reliability of the head-based rendering and other immersion components.
+
 %\section*{Acknowledgements}

### 2009-06-25 trunk/Paper/paper.tex
Added paragraph about component of immersion

@@
 It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
+
+We are considering FOV and the sensor dropouts, or \emph{presence of the augmented imagery}, as the varying components of immersion in our study.
+This is based off of Slater's definition \cite{slater} of immersion as the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as controllable aspect of the technological system.
 

### 2009-06-25 trunk/Paper/paper.tex
changes

@@
 In our experiment we define performance as effectiveness in following Top Hat.
-We measure performance by recording the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
+The dependent variable we measure is the angular distance (in yaw) between the the user's viewing direction and the direction towards the target character.
 This measure ranges between zero and 180 degrees (we used the smaller of the two options, clockwise and counter-clockwise).
@@
 Finally, we do not observe much difference in performance between the medium and large field of view cases.
-Using Tukey's HSD to compare means, the medium and large field of view conditions are in fact not significantly different (p=??).
+Using Tukey's HSD to compare means, the medium and large field of view conditions are in fact not significantly different ($p=0.575$).
 
@@
 We ran a post-hoc analysis to determine which conditions were significantly different.
-Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p=??), but not significantly different from each other (p=??).
+Using Tukey's HSD, we found that the one second and two second conditions were significantly different than zero ($p<0.001$ for both), but not significantly different from each other ($p=0.239$).
 This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
@@
 With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
-%For future work, we could find the threshold between 0.5 and 1.
 
@@
 	\includegraphics[width=2.5in]{figures/extra.pdf}
-	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.
+	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
 We collected this data from seven participants, using the medium field of view.}
@@
 We found that for our task, a low field of view has an adverse effect on performance.  
-This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.
-With reliable head-based rendering, a medium field of view is enough for good performance.
-We did not see a significant performance increase when increasing to a high field of view.
+%This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.
+With reliable head-based rendering, a medium field of view is enough for good performance.  We did not see a significant performance increase when increasing to a high field of view.
 

### 2009-06-25 Paper/paper.tex
Added independent variable notation

@@
 In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
-Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between reliability and field of view in an AR system.
+Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between sensor reliability and field of view in an AR system.
 
@@
 
-In our experiment we vary two parameters: the field of view of the AR interface and the length of sensor dropout periods.
+In our experiment we have two independent variables: the field of view of the AR interface and the length of sensor dropout periods.
 Each trial is 60 seconds, with seven sensor dropouts.
@@
 	\centering
-	\includegraphics[width=3.5in]{figures/extra.pdf}
-	\caption{Box plot of the number of times the participant loses the target versus the length of the sensor dropout periods, including the half second dropout condition.
+	\includegraphics[width=2.5in]{figures/extra.pdf}
+	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
 We collected this data from seven participants, using the medium field of view.}
@@
 
+We designed an experiment to study a target following task which makes use of X-ray vision through an augmented reality device.
+We used an augmented reality simulation to make the experiment feasible.
+
+We found that for our task, a low field of view has an adverse effect on performance.  
+%This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.
+With reliable head-based rendering, a medium field of view is enough for good performance.  We did not see a significant performance increase when increasing to a high field of view.
+
+Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.
+However, for our task, the performance decrease levels off at one second dropouts.
+Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.
+This suggests that below a certain length of dropouts, good level of performance can be maintained, provided that the field of view is large enough.
+This threshold on the dropouts would change depending on what is being followed, but we expect the same pattern would appear in other target following scenarios.
+
+Future work lies in further investigating the effect of the dropout length 
+below the one second level.  Also, this work points to other interesting experiments which might test the interaction between reliability of the head-based rendering and other immersion components.
+
 %\section*{Acknowledgements}

### 2009-06-25 Paper/paper.tex

@@
 	\centering
-	\includegraphics[width=3in]{figures/augmentedroom.png}
+	\includegraphics[width=2.5in]{figures/augmentedroom.png}
 	\caption{Overhead view of virtual room (approx.
@@
 
-We designed an experiment to study a target following task which makes use of X-ray vision through an augmented reality device.
-We used an augmented reality simulation to make the experiment feasible.
+We designed a AR experiment to study the effects of tracker reliability and field of view on a target following task.
+We developed a virtual reality based AR simulation to enable control over these independent variables.
 
@@
 Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.
-However, for our task, the performance decrease levels off at one second dropouts.
+However, the performance decrease levels off at one second dropouts.
 Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.
-This suggests that below a certain length of dropouts, good level of performance can be maintained, provided that the field of view is large enough.
-This threshold on the dropouts would change depending on what is being followed, but we expect the same pattern would appear in other target following scenarios.
+This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.
+This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
 
 Future work lies in further investigating the effect of the dropout length 
-below the one second level.  Also, this work points to other interesting experiments which might test the interaction between reliability of the head-based rendering and other immersion components.
+below the one second level.  Our work also suggests new experiments designed to test the interaction between reliability of the head-based rendering and various other immersion components.
 

### 2009-06-25 trunk/Paper/paper.tex

@@
 Many confederates would be needed to walk around the room, and their movements would need to be accurately reproduced for each new participant.
-Also, the confederates would need to be continously tracked for the augmented overlays to be displayed.
-Instead, we chose to simulate the experiment setup in a completely virtual environment.
+Also, the confederates would need to be continously tracked for the augmented overlays to be displayed.%  Instead, we chose to simulate the experiment setup in a completely virtual environment.%
 
@@
 The participant is centered inside a room with various doors and windows giving a view to the outside world.
-They may change their orientation as the please, but not change position (3 degrees of freedom).
-The virtual person to be tracked walks an unpredictable path outside the room.
-The user's view to the virtual person may be occluded at any time by the walls of the building; at other times the virtual person may be visible through doors or windows in the building.
-The augmented view element is composed of red marker rectangles indicating the location of the tracked person and other virtual people.
-
-Participants are presented a view of the scene through a head-mounted display helmet with orientation tracking.
-The augmented overlays are merged with the original image to form the simulated augmented reality view.
+They may change their orientation, but not change position (3 degrees of freedom).
+The user's view to the virtual people may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
+The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and distractor people.
 
@@
 
-We hypothesize that less immersive conditions will have longer and more frequent periods where the participant loses the target.
-We considered several different metrics which might be appropriate in testing this hypothesis.
+%We hypothesize that less immersive conditions will have longer and more frequent periods where the participant loses the target.
+We considered several different metrics which might be appropriate in testing this hypothesis.%
 

### 2009-06-25 Paper/paper.tex
added text about display options

@@
 For example, such dropouts might occur with a magnetic sensor near interfering material.
+Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.
+However, these options could have a positive or negative effect on performance, depending on the circumstances.
+We determined that hiding the AR overlay during sensor dropouts has the least effect on performance.
 
@@
 
-We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.
+We performed two-way {ANOVA} for the time following target metric as an overall analysis of the effects of our two factors.
 Figure \ref{fig:anova} gives the results of this analysis.
@@
 Figure \ref{fig:interaction} shows a plot of this interaction.
+% TODO: need F values reported here.
 

### 2009-06-25 Paper/paper.tex
removed one line

@@
 Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.
-However, these options could have a positive or negative effect on performance, depending on the circumstances.
-We determined that hiding the AR overlay during sensor dropouts has the least effect on performance.
+%However, these options could have a positive or negative effect on performance, depending on the circumstances.
+However, we determined that hiding the AR overlay during sensor dropouts would have the least effect on performance.
 

### 2009-06-25 trunk/Paper/paper.tex

@@
 Other virtual cars and their associated white boxes acted as distractors.
-At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.
-While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.
+At specific times during the experiment, the simulation would freeze and the participant identified the location of the correct car.
 
@@
 	\centering
-	\includegraphics[width=7in]{figures/triple_view.png}
+	\includegraphics[width=5in]{figures/triple_view.png}
 	\caption{Sample views of AR simulation.
@@
 For example, such dropouts might occur with a magnetic sensor near interfering material.
+Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.
+%However, these options could have a positive or negative effect on performance, depending on the circumstances.
+However, we determined that hiding the AR overlay during sensor dropouts would have the least effect on performance.
 
@@
 Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.
-This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.
-This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
+This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.%  This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
 

### 2009-06-25 trunk/Paper/paper.tex
changed a little more

@@
 From this plot we can see that with a small field of view, performance is equally bad with no dropouts or one second dropouts.  
-The small field of view seems to have a dominating effect on performance.
+%The small field of view seems to have a dominating effect on performance.  
 When the field of view is increased to the medium level, performance dramatically increases when there are no dropouts, but only gets slightly better when there are dropouts.
-Finally, we do not observe much difference in performance between the medium and large field of view cases.
+Finally, there is not a significant difference in performance between the medium and large field of view ($p=0.575$), which suggests that the medium field of view is enough for good performance.
+%Finally, we do not observe much difference in performance between the medium and large field of view cases.
 Using Tukey's HSD to compare means, the medium and large field of view conditions are in fact not significantly different ($p=0.575$).

### 2009-06-25 Paper/paper.tex
abstract

@@
 
-We examine the effects of level of immersion and tracking sensor reliability in augmented reality (AR) X-ray vision with a virtual reality (VR) based simulation.
-We analyze participant performance on a target following task while varying the field of view of the AR display, as well as the reliability of the head tracking sensor.
+We examine the effect of varying levels of immersion on the performance of a target following task in simulated augmented reality (AR) X-ray vision.
+We do this using virtual reality (VR) based simulation.
+We analyze participant performance while varying the field of view of the simulated AR display, as well as the reliability of the head tracking sensor as our factors of immersion.
 In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
-Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between sensor reliability and field of view in an AR system.
+Our study gives insight into the effect of tracking sensor reliability , as well as the relationship between sensor reliability and field of view on user performance in a target following task in a simulated AR system.
 

### 2009-06-25 Paper/paper.tex

@@
 
-We examine the effect of varying levels of immersion on the performance of a target following task in simulated augmented reality (AR) X-ray vision.
+We examine the effect of varying levels of immersion on the performance of a target following task in augmented reality (AR) X-ray vision.
 We do this using virtual reality (VR) based simulation.
-We analyze participant performance while varying the field of view of the simulated AR display, as well as the reliability of the head tracking sensor as our factors of immersion.
+We analyze participant performance while varying the field of view of the AR display, as well as the reliability of the head tracking sensor as our components of immersion.
 In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
-Our study gives insight into the effect of tracking sensor reliability , as well as the relationship between sensor reliability and field of view on user performance in a target following task in a simulated AR system.
+Our study gives insight into the effect of tracking sensor reliability, as well as the relationship between sensor reliability and field of view on user performance in a target following task in a simulated AR system.
 

### 2009-06-25 trunk/Paper/paper.tex
more p values

@@
 
-We examine the effects of level of immersion and tracking sensor reliability in augmented reality (AR) X-ray vision with a virtual reality (VR) based simulation.
-We analyze participant performance on a target following task while varying the field of view of the AR display, as well as the reliability of the head tracking sensor.
+We examine the effect of varying levels of immersion on the performance of a target following task in augmented reality (AR) X-ray vision.
+We do this using virtual reality (VR) based simulation.
+We analyze participant performance while varying the field of view of the AR display, as well as the reliability of the head tracking sensor as our components of immersion.
 In low reliability conditions, we simulate sensor dropouts by disabling the augmented view of the scene for brief time periods.
-Our study gives insight into the effect of tracking sensor reliability on user performance in a target following task, as well as the relationship between sensor reliability and field of view in an AR system.
+Our study gives insight into the effect of tracking sensor reliability, as well as the relationship between sensor reliability and field of view on user performance in a target following task in a simulated AR system.
 
@@
 Figure \ref{fig:anova} gives the results of this analysis.
-Both the failure length and the field of view have a very significant effect on performance using this metric.
+Both the failure length and the field of view have a significant effect on performance with this metric, at the $p=0.001$ level.
 % TODO: interaction?
@@
 We also used a two-way {ANOVA} to examine the number of times the participant loses the target, with results shown in Figure \ref{fig:anova}.
-Both field of view and dropout length have a significant effect.
-There also is an interaction between the two factors.
+Both field of view and dropout length have a significant effect at the $p=0.001$ level.
+There also is an interaction between the two factors at the $p=0.05$ level.
 Figure \ref{fig:interaction} shows a plot of this interaction.

### 2009-06-25 Paper/paper.tex
fixed conclusions

@@
 
-We found that for our task, a low field of view has an adverse effect on performance.  
+We found that for our task, good performance could not be achieved with a low field of view, even with perfect head-based rendering reliability.
+%has an adverse effect on performance.   even when introducing medium-length sensor dropouts
 %This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.

### 2009-06-25 trunk/Paper/paper.tex
added umlaut

@@
 
-Bane and Hollerer propose a set of AR tools for virtual X-ray vision \cite{1383060}.
+Bane and H\"{o}llerer propose a set of AR tools for virtual X-ray vision \cite{1383060}.
 Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.

### 2009-06-25 trunk/Paper/paper.tex

@@
 Unfortunately, implementing X-ray vision is not a trivial task.
-AR systems certain unsolved issues, including those pertaining to tracking the objects in the X-ray view.
+AR systems face certain unresolved issues, including those pertaining to tracking objects in the X-ray view.
 For example, GPS dropout errors \cite{4079263} may cause augmented position information to be temporarily lost.
@@
 A task common to many AR X-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
-Problems with the AR system tracking objects accurately is often an issue which is likely to have an effect on task performance.
-Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the augmented field of view s well as its interaction with tracking difficulties.
-We created a task that aimed to accomplish these goals in a generic fashion in order to be generalizable while still being grounded in reality.
+Problems with the AR system tracking objects accurately is an issue which is likely to have an effect on task performance.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the varied size of augmented field of view as well as its interaction with tracking difficulties.
+We created a task that aimed to study these effect in a generic fashion in order to be generalizable while still being grounded in reality.
 
-In this study, we have implemented an AR simulation with virtual reality hardware and software to examine the effects of varying FOV and sensor dropouts in a target following task.
-Using an AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
+In this study, we have implemented an AR simulation in virtual reality to examine the effects of varying FOV and sensor dropouts in a target following task.
+An AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
 It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
 
-We are considering FOV and the sensor dropouts, or \emph{presence of the augmented imagery}, as the varying components of immersion in our study.
+We are considering FOV and the sensor dropouts, or persistence of the augmented imagery, as the varying components of immersion in our study.
 This is based off of Slater's definition \cite{slater} of immersion as the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as controllable aspect of the technological system.
@@
 Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own \cite{4637329}.
-Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
+Participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.
 A white box was continuously visible, even when the car itself was occluded by a building in the environment.
@@
 
-Ragan et.
-al.
-\cite{4811058} developed an AR simulation using virtual reality to examine the effect of registration errors on an object manipulation tasks.
+Ragan et al.
+developed an AR simulation using virtual reality to examine the effect of registration errors on an object manipulation tasks \cite{4811058}.
 The implemented task involved moving a virtual ring from one end of an irregularly shaped tube to the other, while avoiding collisions between the two objects.
@@
 The user's view to the virtual people may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
-The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and distractor people.
+The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and the distractor people.
 
@@
 %However, these options could have a positive or negative effect on performance, depending on the circumstances.
-However, we determined that hiding the AR overlay during sensor dropouts would have the least effect on performance.
+However, we determined through expert analysis that hiding the AR overlay during sensor dropouts would have the least effect on performance.
 
-In our experiment we have two independent variables: the field of view of the AR interface and the length of sensor dropout periods.
+The experiment has two independent variables: the field of view of the AR interface and the length of sensor dropout periods.
 Each trial is 60 seconds, with seven sensor dropouts.
@@
 We experimented with longer dropout periods but deemed them too long.
-This gave us a total of 9 conditions, and for each participant we tested each condition 3 times giving a total of 27 trials per participant.
+There is a total of 9 conditions.
+For each participant we tested each condition 3 times giving a total of 27 trials per participant.
 

### 2009-08-24 Paper/paper.tex
added author list

@@
 \author{
-Submission 165
-%Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ Starbucks Research %
-%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com}\\Nigel Mansell\thanks{nigelf1@msn.com}\\ Grimley Widgets, Inc. %
-%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}\\ Martha Stewart Enterprises \\ Microsoft Research
+\begin{tabular}{cccc}
+Jonathan Ventura\hspace{12pt}
+Marcus Jang\hspace{12pt}
+Tyler Crain\hspace{12pt}
+Tobias H\"{o}llerer\hspace{12pt}
+\end{tabular}\\University of California, Santa Barbara
+\and Doug Bowman\\Virginia Tech
 }
@@
 %%%%%% START OF THE PAPER %%%%%%
-
+%
+\def\sharedaffiliation{%
+\end{tabular}
+\begin{tabular}{c}}
+%
 \begin{document}

### 2009-08-24 trunk/Paper/paper.tex
first page revisions

@@
 A task common to many AR X-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
-Problems with the AR system tracking objects accurately is an issue which is likely to have an effect on task performance.
-Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the varied size of augmented field of view as well as its interaction with tracking difficulties.
-We created a task that aimed to study these effect in a generic fashion in order to be generalizable while still being grounded in reality.
+Accuracy in AR system object tracking is an issue which is likely to have an effect on task performance.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the varying sizes of augmented field of view as well as its interaction with tracking imperfections.
+We created a task that aimed to study these effects in a generic fashion in order to be generalizable while still being grounded in reality.
 
 In this study, we have implemented an AR simulation in virtual reality to examine the effects of varying FOV and sensor dropouts in a target following task.
-An AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
-It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
+We simulate video see-through AR using a VR head mounted display.
+This AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
+It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibility changes, weather problems, or environmental distractions.
 
 We are considering FOV and the sensor dropouts, or persistence of the augmented imagery, as the varying components of immersion in our study.
-This is based off of Slater's definition \cite{slater} of immersion as the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as controllable aspect of the technological system.
+We adhere to Slater's definition \cite{slater} of immersion as being the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as the directly controllable aspects of the technological system.
 
-The AR simulation is built as a target following task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
-We wish to answer the following questions.
+The AR simulation is built as a target following task.
+The participant stands inside a virtual room and visually follows a person walking an unpredictable path outside the room.
+We wish to answer the subsequent questions.
 What effect does varying the augmented FOV have on user task performance?
@@
 Bane and H\"{o}llerer propose a set of AR tools for virtual X-ray vision \cite{1383060}.
-Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.
-The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.
+Various augmented views are presented, each attempting to overcome "Superman's X-ray vision" problem of presenting too much AR information to the user.
+The toolset's usefulness is illustrated with an example of an outdoor user viewing the contents of a nearby building.
 

### 2009-08-25 NewResults/numtimeslost_factor.py

@@ -40,6 +40,9 @@ while not done:
 
 		#print "%d samples"%len(samples_in)
 
+		if fov != 'lowfov':
+		    continue
+
 		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))
 		totaltime = 0
 		

### 2009-08-25 trunk/Paper/paper.tex
minor changes through beginning of page 3

@@
 Our experiment involves a target following task which uses an augmented reality interface to allow people to be tracked behind walls.
-A participant stands in the middle of a square room with doors and windows on all sides.
+A participant stands in the middle of a square room with unique doors and/or windows on each wall.
 Virtual people walk unpredictable paths outside the building.
@@
 
-The design of the experiment is built with several goals in mind.
+The experiment was designed with several goals in mind.
 The task should have straightforward, quantitative results.
 Also, the task should be reasonably simple so as not to frustrate the participants.
-Further, the task should be generic so as to be generizable, yet grounded in reality (i.e.
+Furthermore, the task should be generic so as to be generizable, yet grounded in reality (i.e.
 not an abstract world), to enhance participant familiarity with the AR task.
@@
 They may change their orientation, but not change position (3 degrees of freedom).
-The user's view to the virtual people may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
+The user's view to the virtual person may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
 The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and the distractor people.
@@
 We introduce periods of sensor dropouts, where the augmented overlays disappear, to simulate the effect of dropouts in a real tracking system.
-For example, such dropouts might occur with a magnetic sensor near interfering material.
+Such dropouts may occur with a magnetic sensor near interfering material.
 Different display options are available in this case; for example we could use the last known sensor reading, or predict future values.
@@
 The experiment has two independent variables: the field of view of the AR interface and the length of sensor dropout periods.
-Each trial is 60 seconds, with seven sensor dropouts.
-The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-The length of dropouts varied between two seconds (highest), one second (medium) and zero seconds (lowest).
+Each trial lasts 60 seconds, and includes seven sensor dropouts.
+The total vertical field of view of the HMD is 36 degrees, while the three possible values for the augmented field of view are 10, 20, and 34 degrees.
+The length of dropouts vary between two seconds (highest), one second (medium) and zero seconds (lowest).
 We experimented with longer dropout periods but deemed them too long.
@@
 
-During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat walking an unpredictable path outside of the room.
+During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat; each walking an unpredictable path outside of the room.
 The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
-Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
-At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
-Top Hat had the additional constraints as follows.
-He must start at the same location for each trial which made him visible by the participant through the front door the building.
-In order to make the paths by of similar difficulty Top Hat must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
+Paths is constrained to stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
+At each path point, the virtual person randomly changes its walking speed within the range of 4 to 7 \emph{m/s}.
+Top Hat has the additional constraints as follows.
+He starts each trial just outside the front door of the building, standing in the participant's plain view.
+In order to make the paths by of similar difficulty Top Hat walks around the entire building at least once, and is visible through the windows for at least 10 seconds.
 These constraints were met by randomly generating paths until all were satisfied.
-We generated 27 sets of paths in this fashion, each corresponding to a specific trial.
-We used a latin squares ordering to discourage the effects of learning.
+We generated 27 sets of paths in this fashion, each corresponding to a specific trial and used a latin squares ordering to discourage learning effects.
 
@@
 We implemented this experiment with Python and the Vizard virtual reality toolkit.
-Head-tracking was handled by a InterSense InertiaCube3 orientation tracker with a 180 Hz refresh rate.
+Head-tracking was handled with a InterSense InertiaCube3 orientation tracker with a refresh rate of 180 Hz.
 We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
-As we were not concerned the effects of stereo, we turned off the stereoscopic feature.
-The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
+As we were not concerned the effects of stereo, we disabled the stereoscopic feature.
+The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 GB of RAM, a NVIDIA GeForce 9800 GX2 graphics card and Windows XP SP3.
 
@@
 
-We ran this experiment on 19 subjects, 11 were male, and 8 female.
+We ran this experiment on 19 participants; 11 male, and 8 female.
 The participants ranged in age from 23 to 59, 15 were in the range 23 to 30, and the remaining 4 in the range 51 to 49.
 None were colorblind.
-Most users were heavy computer users with at least somewhat familiarity with virtual reality and augmented reality.
-Only one participant played 3D video games often.
-After every 9 trials we had the participants take a mandatory 5 minute break.
-Some experienced light fatigue, but none experienced any strong effects of motion sickness.
+Most users were heavy computer users with at least some familiarity with virtual reality and augmented reality.
+Only one participant reported regularly playing 3D video games.
+After every 9 trials, the participants took a mandatory 5 minute break.
+Some experienced light fatigue, but none experienced strong effects of motion sickness.
 
@@
 
-Because the video does not always run at precisely 60 Hz, each trial might have a slightly different number of measurements.
+Because the video does not always run precisely at 60 Hz, each trial might have a slightly different number of measurements.
 We also record a timestamp for each measurement, so that we can determine their exact frequency.

### 2009-08-25 trunk/Paper/paper.tex
finished my minor revisions

@@
 In the target following state, the angular error is generally low.
-We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.
+We had asked participants to keep the cross hairs on the target as close as possible, but different participants achieved different levels of accuracy.
 In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.
@@
 
-We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.
-The threshold specifies the maximum angular error that still represents successful following of the target.
+We may detect the participant's state more explicitly by applying a threshold $\tau$ to the data.
+The threshold specifies the maximum angular error that represents successful following of the target.
 %When the error rises above the threshold, we consider the participant lost.
-While conducting the trials we noticed that the participants spent a majority of the time correctly following Top Hat, computing the median angular error at each time-step across all trials gave us values mostly around 4 degrees so we consider values close to this as successful tracking.
-We also noticed that during trials as Top Hat would change direction the participant's angular error would increase a small amount for a short time, without losing the target.
+When conducting the trials we noticed that the participants spent a majority of the time correctly following Top Hat.
+By computing the median angular error, we found that the participants successfully tracked Top Hat within an error margin of 4 degrees.
+We also noticed that during trials as Top Hat would change direction the participant's angular error would increase by a small amount for a short period of time, without losing the target.
 As a result we doubled the 4 degrees, giving us a threshold of $\tau = 8$ degrees

### 2009-08-25 trunk/Paper/paper.tex
more small changes JDV

@@
 Accuracy in AR system object tracking is an issue which is likely to have an effect on task performance.
-Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the varying sizes of augmented field of view as well as its interaction with tracking imperfections.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of varying the augmented field of view as well as its interaction with tracking imperfections.
 We created a task that aimed to study these effects in a generic fashion in order to be generalizable while still being grounded in reality.
@@
 This AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
-It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibility changes, weather problems, or environmental distractions.
+It also gives us the ability to keep unnecessary factors from generating noise in our data, such as illumination or visibility changes, weather problems, or environmental distractions.
 
@@
 During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat; each walking an unpredictable path outside of the room.
-The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
-Paths is constrained to stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
+The paths for each virtual person were randomly generated from a set of coordinates exterior to the room.
+Paths were constrained to stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
 At each path point, the virtual person randomly changes its walking speed within the range of 4 to 7 \emph{m/s}.
@@
 He starts each trial just outside the front door of the building, standing in the participant's plain view.
-In order to make the paths by of similar difficulty Top Hat walks around the entire building at least once, and is visible through the windows for at least 10 seconds.
+In order to make paths of similar difficulty Top Hat walks around the entire building at least once, and is visible through the windows for at least 10 seconds.
 These constraints were met by randomly generating paths until all were satisfied.
@@
 In the target following state, the angular error is generally low.
-We had asked participants to keep the cross hairs on the target as close as possible, but different participants achieved different levels of accuracy.
+We had asked participants to keep the cross hairs on the target as closely as possible, but different participants achieved different levels of accuracy.
 In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.
@@
 
-The simplest is the average error over an entire trial.
+The simplest metric is the average error over an entire trial.
 This may not accurately represent sensor dropouts, because the error does not necessarily stay high during the lost state, and also different participants may generally keep the cross hair further from the target even in the target following state.
@@
 We also noticed that during trials as Top Hat would change direction the participant's angular error would increase by a small amount for a short period of time, without losing the target.
-As a result we doubled the 4 degrees, giving us a threshold of $\tau = 8$ degrees
+As a result we doubled the 4 degrees, giving us a threshold of $\tau = 8$ degrees.
 % TODO: how to motivate this threshold?
@@
 \end{figure}
+
 Using this threshold we measure a participant's \emph{time to failure}, the time until the a lost state is encountered.

### 2009-08-25 NewResults/mean.R
R changes for paper

@@ -1,4 +1,4 @@
-datafilename="mean_factor.csv"
+datafilename="mean.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 #data.all[,'mean'] = rowMeans(data.all[,4:3603])
@@ -7,3 +7,9 @@ attach(data.all)
 data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
 colnames(data.agg) <- c("id","fov","deadlen",'mean')
 detach(data.all)
+
+aov.two = aov(numtimes~deadlen*fov+Error(id/(deadlen*fov)),data.agg)
+
+attach(data.agg)
+interaction.plot(fov,deadlen,numtimes)
+detach(data.agg)

### 2009-08-25 NewResults/timetracking_factor.py

@@ -8,14 +8,15 @@ if len(sys.argv) == 2:
 id_dict = {}
 for i in xrange(27):
     id_dict[ "%d"%i ] = "id%d"%i
-fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
-len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+fov_dict = { '10':'10deg', '20':'20deg', '34':'34deg' }
+len_dict = { '0':'0.0s', '0.5':'0.5s', '1':'1.0s', '2':'2.0s' }
 
-resamples = open("resampled.csv")
+#resamples = open("resampled.csv")
 
 while not done:
     for i in xrange(27):
-	line =  resamples.readline()
+	#line =  resamples.readline()
+	line = sys.stdin.readline()
 	if line == "":
 	    done = True
 	    break
@@ -41,4 +42,4 @@ while not done:
 	for j in xrange(60*60):
 	    if samples_in[j] <= thresh:
 		totaltime += 1
-	sys.stdout.write(" %d\n"%totaltime)
+	sys.stdout.write("%f\n"%(float(totaltime)/60.))

### 2009-08-25 trunk/NewResults/numtimeslost_factor.py
R changes for paper

@@ -3,23 +3,28 @@ import sys
 done = False
 
 thresh = 12
-secondsLoss = 2
+secondsLoss = 3
 
 if len(sys.argv) == 2:
     thresh = int(sys.argv[1])
 
+if len(sys.argv) == 3:
+    thresh = int(sys.argv[1])
+    secondsLoss = float(sys.argv[2])
+
 id_dict = {}
 for i in xrange(27):
     id_dict[ "%d"%i ] = "id%d"%i
-fov_dict = { '10':'lowfov', '20':'medfov', '34':'uppfov' }
-len_dict = { '0':'lowlen', '1':'medlen', '2':'upplen' }
+fov_dict = { '10':'10deg', '20':'20deg', '34':'34deg' }
+len_dict = { '0':'0.0s', '0.5':'0.5s', '1':'1.0s', '2':'2.0s' }
 
-resamples = open("resampled.csv")
+#resamples = open("resampled.csv")
 
 while not done:
     for i in xrange(27):
 		
-		line = resamples.readline()
+		#line = resamples.readline()
+		line = sys.stdin.readline()
 		if line == "":
 			done = True
 			break

### 2009-08-25 trunk/NewResults/tt.R

@@ -1,45 +1,18 @@
-datafilename="timetracking_5_factor.csv"
+#datafilename="extratt8.csv"
+datafilename="tt8.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
-
-#data.all[0,1] = 
-#data.all[0,2] = 'fov'
-#data.all[0,3] = 'deadlen'
-#data.all[,'mean'] = rowMeans(data.all[,4:3603])
-
-#plot(data.all[,'avg'],type='lines')
 
 attach(data.all)
 data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
 colnames(data.agg) <- c("id","fov","deadlen","tt")
 
-#data.agg[sapply(data.agg['fov'],function(x) x==10),'fov'] = 'lowfov'
-#data.agg[sapply(data.agg['fov'],function(x) x==20),'fov'] = 'medfov'
-#data.agg[sapply(data.agg['fov'],function(x) x==34),'fov'] = 'uppfov'
-#
-#data.agg[sapply(data.agg['deadlen'],function(x) x==0),'deadlen'] = 'lowlen'
-#data.agg[sapply(data.agg['deadlen'],function(x) x==1),'deadlen'] = 'medlen'
-#data.agg[sapply(data.agg['deadlen'],function(x) x==2),'deadlen'] = 'upplen'
-#
-#data.agg[sapply(data.agg['id'],function(x) x==0),'id'] = 'id0'
-#data.agg[sapply(data.agg['id'],function(x) x==1),'id'] = 'id1'
-#data.agg[sapply(data.agg['id'],function(x) x==2),'id'] = 'id2'
-#data.agg[sapply(data.agg['id'],function(x) x==3),'id'] = 'id3'
-#data.agg[sapply(data.agg['id'],function(x) x==4),'id'] = 'id4'
-#data.agg[sapply(data.agg['id'],function(x) x==5),'id'] = 'id5'
-#data.agg[sapply(data.agg['id'],function(x) x==6),'id'] = 'id6'
-#data.agg[sapply(data.agg['id'],function(x) x==7),'id'] = 'id7'
-#data.agg[sapply(data.agg['id'],function(x) x==8),'id'] = 'id8'
-#data.agg[sapply(data.agg['id'],function(x) x==9),'id'] = 'id9'
-
 detach(data.all)
 
[…]

### 2009-08-25 NewResults/numtimes.R
adding numtimes file

@@ -1,0 +1,25 @@
+#datafilename="numtimes_extra8_4.csv"
+datafilename="numtimes8_4.csv"
+data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
+
+attach(data.all)
+data.agg = aggregate(V4,list(V1,V2,V3), function(x)mean(x))
+colnames(data.agg) <- c("id","fov","deadlen",'numtimes')
+detach(data.all)
+
+#aov.deadlen = aov(numtimes~deadlen+Error(id/deadlen),data.agg)
+#aov.fov = aov(numtimes~deadlen+Error(id/fov),data.agg)
+aov.two = aov(numtimes~deadlen*fov+Error(id/(deadlen*fov)),data.agg)
+
+#plot(numtimes~deadlen,data.agg,xlab='Dropout Length',ylab='Number of Times Lost')
+#plot(numtimes~fov,data.agg,xlab='Field of View',ylab='Number of Times Lost')
+
+par(mar=c(5,5,5,2),cex.axis=2,cex.lab=2)
+
+
+#attach(data.agg)
+#interaction.plot(fov,deadlen,numtimes,xlab='Field of View',ylab='Number of Times Lost',trace.label='Dropout Len')
+#detach(data.agg)
+
+plot(numtimes~deadlen,data.agg,cex.axis=2,cex.lab=2,xlab='Dropout Length',ylab='Number of Times Lost')
+#plot(numtimes~fov,data.agg,cex.axis=2,cex.lab=2,xlab='Field of View',ylab='Number of Times Lost')

### 2009-08-25 NewResults/numtimeslost_factor.py
updating numtimes scripts

@@ -4,6 +4,7 @@ done = False
 
 thresh = 12
 secondsLoss = 3
+selectedFov = '10deg'
 
 if len(sys.argv) == 2:
     thresh = int(sys.argv[1])
@@ -11,6 +12,12 @@ if len(sys.argv) == 2:
 if len(sys.argv) == 3:
     thresh = int(sys.argv[1])
     secondsLoss = float(sys.argv[2])
+
+if len(sys.argv) == 4:
+    thresh = int(sys.argv[1])
+    secondsLoss = float(sys.argv[2])
+    selectedFov = sys.argv[3]
+
 
 id_dict = {}
 for i in xrange(27):
@@ -45,7 +52,7 @@ while not done:
 
 		#print "%d samples"%len(samples_in)
 
-		if fov != 'lowfov':
+		if fov != selectedFov:
 		    continue
 
 		sys.stdout.write("%s,%s,%s,"%(id,fov,deadlen))

### 2009-08-25 trunk/NewResults/numtimes.R
updating numtimes scripts

@@ -1,5 +1,5 @@
 #datafilename="numtimes_extra8_4.csv"
-datafilename="numtimes8_4.csv"
+datafilename="numtimeslost8_4_34.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 attach(data.all)
@@ -9,7 +9,7 @@ detach(data.all)
 
 #aov.deadlen = aov(numtimes~deadlen+Error(id/deadlen),data.agg)
 #aov.fov = aov(numtimes~deadlen+Error(id/fov),data.agg)
-aov.two = aov(numtimes~deadlen*fov+Error(id/(deadlen*fov)),data.agg)
+#aov.two = aov(numtimes~deadlen*fov+Error(id/(deadlen*fov)),data.agg)
 
 #plot(numtimes~deadlen,data.agg,xlab='Dropout Length',ylab='Number of Times Lost')
 #plot(numtimes~fov,data.agg,xlab='Field of View',ylab='Number of Times Lost')

### 2009-08-25 trunk/Paper/paper.tex
changed interaction figure

@@
 \usepackage{graphicx}
+
+\usepackage{subfigure}
 
@@
 \keywords{
-%radiosity, global illumination, constant time
+augmented reality, immersion, user study, simulation 
 }
@@
 %%%%%% START OF THE PAPER %%%%%%
-%
-\def\sharedaffiliation{%
-\end{tabular}
-\begin{tabular}{c}}
-%
 \begin{document}
@@
 \begin{CRcatlist}
-%  \CRcat{K.6.1}{Management of Computing and Information Systems}{Project and People Management}{Life Cycle};
-%  \CRcat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
+  \CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Augmented Reality};
+  \CRcat{H.1.2}{Models and Principles}{User/Machine Systems}{Human Factors};
 \end{CRcatlist}
@@
 
-%% \copyrightspace
+\copyrightspace
 
@@
 There also is an interaction between the two factors at the $p=0.05$ level.
-Figure \ref{fig:interaction} shows a plot of this interaction.
+Figures \ref{subfig:low}-(c) show box plots for these conditions.
 % TODO: need F values reported here.
@@
 
-\begin{figure}[t]
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=2.5in]{figures/numtimes_interaction.pdf}
+%	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
+%\end{figure}
+
+%\begin{figure*}[t]
+%	\centering
+%	\subfigure[]{\label{subfig:low}\includegraphics[width=1.5in]{figures/numtimes_deadlen_lowfov.pdf}}
+%	\subfigure[]{\label{subfig:med}\includegraphics[width=1.5in]{figures/numtimes_deadlen_medfov.pdf}}
+%	\subfigure[]{\label{subfig:upp}\includegraphics[width=1.5in]{figures/numtimes_deadlen_uppfov.pdf}}
+%	\caption{\label{fig:interaction}Box plots relating sensor dropout length to the number of times the participant loses the target, with (a) low, (b) medium and (c) upper field of view.}
+%\end{figure*}
+
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=1.5in]{figures/extra.pdf}
+%	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected this data from seven participants, using the medium field of view.}
+%\end{figure}
+
+\begin{figure*}[t]
 	\centering
-	\includegraphics[width=2.5in]{figures/numtimes_interaction.pdf}
-	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
-\end{figure}
+	\subfigure[]{\label{subfig:low}\includegraphics[width=1.7in]{figures/numtimes_deadlen_lowfov.pdf}}
+	\subfigure[]{\label{subfig:med}\includegraphics[width=1.7in]{figures/numtimes_deadlen_medfov.pdf}}
+	\subfigure[]{\label{subfig:upp}\includegraphics[width=1.7in]{figures/numtimes_deadlen_uppfov.pdf}}
+	\subfigure[]{\label{subfig:extra}\includegraphics[width=1.7in]{figures/extra.pdf}}
+	\caption{\label{fig:interaction}Box plots (a)-(c) relate sensor dropout length to the number of times the participant loses the target, with (a) low, (b) medium and (c) upper field of view.
+Box plot (d) shows the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected the data for box plot (d) from seven participants, using the medium field of view.}
+\end{figure*}
+
+Figure \ref{fig:tt_deadlen} shows a box plot of time following target versus dropout length.
+Performance decreases between zero and one seconds, but seems to level off between one and two.
+We ran a post-hoc analysis to determine which conditions were significantly different.
+Using Tukey's HSD, we found that the one second and two second conditions were significantly different than zero ($p<0.001$ for both), but not significantly different from each other ($p=0.239$).
+This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
+% TODO: table with values
 
@@
 
-Figure \ref{fig:tt_deadlen} shows a box plot of time following target versus dropout length.
-Performance decreases between zero and one seconds, but seems to level off between one and two.
-We ran a post-hoc analysis to determine which conditions were significantly different.
-Using Tukey's HSD, we found that the one second and two second conditions were significantly different than zero ($p<0.001$ for both), but not significantly different from each other ($p=0.239$).
-This suggests that we may be seeing a thresholding effect above one second sensor dropouts.
-% TODO: table with values
-
 To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
-Figure \ref{fig:extra} shows a box plot of time following target versus sensor dropout length for these seven participants.
+Figure \ref{subfig:extra} shows a box plot of time following target versus sensor dropout length for these seven participants.
 Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.
@@
 With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
-
-\begin{figure}[htb]
-	\centering
-	\includegraphics[width=2.5in]{figures/extra.pdf}
-	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
-We collected this data from seven participants, using the medium field of view.}
-\end{figure}
 
@@
 
-%\section*{Acknowledgements}
-%Cha Lee, Steffen Gauglitz, funding, and all of our participants.
+\section*{Acknowledgements}
+We would like to thank Cha Lee and Steffen Gauglitz for help in designing and running our experiments, and our study participants.
+This work was partially supported by NSF CAREER grant IIS-0747520, and NSF IGERT grant DGE-0221713.
 
@@
 \bibliography{paper}
+
 \end{document}

### 2009-08-25 NewResults/numtimes.R

@@ -1,5 +1,5 @@
 #datafilename="numtimes_extra8_4.csv"
-datafilename="numtimeslost8_4_34.csv"
+datafilename="numtimeslost8_4_10.csv"
 data.all=read.csv(datafilename,header=FALSE)	# read the data into a table
 
 attach(data.all)
@@ -21,5 +21,5 @@ par(mar=c(5,5,5,2),cex.axis=2,cex.lab=2)
 #interaction.plot(fov,deadlen,numtimes,xlab='Field of View',ylab='Number of Times Lost',trace.label='Dropout Len')
 #detach(data.agg)
 
-plot(numtimes~deadlen,data.agg,cex.axis=2,cex.lab=2,xlab='Dropout Length',ylab='Number of Times Lost')
+plot(numtimes~deadlen,data.agg,cex.axis=2,cex.lab=2,xlab='Dropout Length',ylab='Number of Times Lost',ylim=c(0,2.5))
 #plot(numtimes~fov,data.agg,cex.axis=2,cex.lab=2,xlab='Field of View',ylab='Number of Times Lost')

### 2009-08-25 Results/script.R

@@ -52,19 +52,19 @@ detach(data.all)
 
 
 # one-way repeated measures anova on deadlen
-aov.deadlen = aov(mean ~deadlen+Error(id/deadlen),data.agg)
-print(summary(aov.deadlen))
-boxplot(mean~deadlen,data=data.agg)
+#aov.deadlen = aov(mean ~deadlen+Error(id/deadlen),data.agg)
+#print(summary(aov.deadlen))
+#boxplot(mean~deadlen,data=data.agg)
 
 
 
 # two-way anova repeated measures on both
-#aov.two = aov(mean ~(fov*deadlen)+Error(id/(fov*deadlen)),data.agg)
-#print(summary(aov.two))
+aov.two = aov(mean ~(fov*deadlen)+Error(id/(fov*deadlen)),data.agg)
+print(summary(aov.two))
 #boxplot(mean ~fov*deadlen,data=data.agg)
-#attach(data.agg)
-#interaction.plot(fov,deadlen, mean)    #another way to graph the interaction
-#detach(data.agg)
+attach(data.agg)
+interaction.plot(fov,deadlen, mean)    #another way to graph the interaction
+detach(data.agg)
 
 
 

### 2009-08-25 Paper/paper.tex
dougs changes

@@
 \begin{tabular}{cccc}
-Jonathan Ventura\hspace{12pt}
+Jonathan Ventura\thanks{e-mail: jventura@cs.ucsb.edu}\hspace{12pt}
 Marcus Jang\hspace{12pt}
@@
 
-%% \copyrightspace
+\copyrightspace
 
 Augmented Reality systems may be used to simulate X-ray vision.
-The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality}, medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.
+The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality} to medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.
 Unfortunately, implementing X-ray vision is not a trivial task.
@@
 A task common to many AR X-ray systems is to follow, using the augmented overlay, a moving target that is occluded by objects in the real world.
-Problems with the AR system tracking objects accurately is an issue which is likely to have an effect on task performance.
-Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of the varied size of augmented field of view as well as its interaction with tracking difficulties.
-We created a task that aimed to study these effect in a generic fashion in order to be generalizable while still being grounded in reality.
+Accuracy in AR system object tracking is an issue which is likely to have an effect on task performance.
+Since the augmented view is overlaid on top of the real view it is also interesting to study the effects of varying the augmented field of view as well as its interaction with tracking imperfections.
+We created a task that aimed to study these effects in a generic fashion in order to be generalizable while still being grounded in reality.
 
 In this study, we have implemented an AR simulation in virtual reality to examine the effects of varying FOV and sensor dropouts in a target following task.
-An AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
-It also gives us the ability to remove unnecessary factors from generating noise in our data, such as illumination or visibilty changes, weather problems, or environmental distractions.
+We simulate video see-through AR using a VR head mounted display.
+This AR simulation gives us the ability to vary experiment conditions which may be difficult or even impossible to replicate on a real-world AR system.
+It also gives us the ability to keep unnecessary factors from generating noise in our data, such as illumination or visibility changes, weather problems, or environmental distractions.
 
 We are considering FOV and the sensor dropouts, or persistence of the augmented imagery, as the varying components of immersion in our study.
-This is based off of Slater's definition \cite{slater} of immersion as the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as controllable aspect of the technological system.
+We adhere to Slater's definition \cite{slater} of immersion as being the ``objective level of fidelity of the sensory stimuli produced by a technological system,''  enabling us to view the components of immersion as the directly controllable aspects of the technological system.
 
-The AR simulation is built as a target following task, where the participant is placed inside a virtual room and visually follows a person outside, walking an unpredictable path.
-We wish to answer the following questions.
+The AR simulation is built as a target following task.
+The participant stands inside a virtual room and visually follows a person walking an unpredictable path outside the room.
+We wish to answer the subsequent questions.
 What effect does varying the augmented FOV have on user task performance?
@@
 The participant is centered inside a room with various doors and windows giving a view to the outside world.
-They may change their orientation, but not change position (3 degrees of freedom).
-The user's view to the virtual people may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
+They may change their orientation, but not change position (three degrees of freedom).
+The user's view to the virtual person may be occluded at any time by the walls of the building; at other times the person may be visible through doors or windows in the building.
 The augmented view element is composed of red marker rectangles indicating the location of the person to be followed and the distractor people.
@@
 The experiment has two independent variables: the field of view of the AR interface and the length of sensor dropout periods.
-Each trial is 60 seconds, with seven sensor dropouts.
-The total vertical field of view of the HMD is 36 degrees; the three possible values for the augmented field of view were 10, 20, and 34 degrees.
-The length of dropouts varied between two seconds (highest), one second (medium) and zero seconds (lowest).
+Each trial lasts 60 seconds, and includes seven sensor dropouts.
+The total vertical field of view of the HMD is 36 degrees, while the three possible values for the augmented field of view are 10, 20, and 34 degrees.
+The length of dropouts vary between two seconds (highest), one second (medium) and zero seconds (lowest).
 We experimented with longer dropout periods but deemed them too long.
-There is a total of 9 conditions.
-For each participant we tested each condition 3 times giving a total of 27 trials per participant.
+There is a total of nine conditions.
+For each participant we tested each condition three times giving a total of 27 trials per participant.
 
-During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat walking an unpredictable path outside of the room.
-The paths for each virtual person was randomly generated from set of coordinates exterior to the room.
-Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
-At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.
-Top Hat had the additional constraints as follows.
-He must start at the same location for each trial which made him visible by the participant through the front door the building.
-In order to make the paths by of similar difficulty Top Hat must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.
+During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat; each walking an unpredictable path outside of the room.
+The paths for each virtual person were randomly generated from a set of coordinates exterior to the room.
+Paths were constrained to stay within the rectangular $50\times50$ $m^2$ area surrounding the room.
+At each path point, the virtual person randomly changes its walking speed within the range of four to seven meters per second.
+%\emph{m/s}.
+Top Hat has the additional constraints as follows.
+He starts each trial just outside the front door of the building, standing in the participant's plain view.
+In order to make paths of similar difficulty Top Hat walks around the entire building at least once, and is visible through the windows for at least 10 seconds.
 These constraints were met by randomly generating paths until all were satisfied.
-We generated 27 sets of paths in this fashion, each corresponding to a specific trial.
-We used a latin squares ordering to discourage the effects of learning.
+We generated 27 sets of paths in this fashion, each corresponding to a specific trial and used a latin squares ordering to discourage learning effects.
 
@@
 	\centering
-	\includegraphics[width=2.5in]{figures/augmentedroom.png}
+	\includegraphics[width=2.15in]{figures/augmentedroom.png}
 	\caption{Overhead view of virtual room (approx.
@@
 We implemented this experiment with Python and the Vizard virtual reality toolkit.
-Head-tracking was handled by a InterSense InertiaCube3 orientation tracker with a 180 Hz refresh rate.
+Head-tracking was handled with a InterSense InertiaCube3 orientation tracker with a refresh rate of 180 Hz.
 We used a Pro-View 60 head mounted display (HMD) displaying $640\times480$ video at 60 Hz, and 36 degrees vertical field of view.
-As we were not concerned the effects of stereo, we turned off the stereoscopic feature.
-The experiment ran on a 2.4 GHz Intel Core 2 machine with 2 Gb of RAM and a NVIDIA GeForce 9800 GX2 graphics card running Windows XP SP3.
+As we were not concerned with the effects of stereo, we disabled the stereoscopic feature.
+The experiment ran on a 2.4 GHz Intel Core 2 machine with two gigabytes of RAM, a NVIDIA GeForce 9800 GX2 graphics card and Windows XP SP3.
 
@@
 
-We ran this experiment on 19 subjects, 11 were male, and 8 female.
-The participants ranged in age from 23 to 59, 15 were in the range 23 to 30, and the remaining 4 in the range 51 to 49.
+We ran this experiment with 19 participants; eleven male and eight female.
+The participants ranged in age from 23 to 59, 15 were in the range 23 to 30, and the remaining four in the range 51 to 49.
 None were colorblind.
-Most users were heavy computer users with at least somewhat familiarity with virtual reality and augmented reality.
-Only one participant played 3D video games often.
-After every 9 trials we had the participants take a mandatory 5 minute break.
-Some experienced light fatigue, but none experienced any strong effects of motion sickness.
+Most users were heavy computer users with at least some familiarity with virtual reality and augmented reality.
+Only one participant reported regularly playing 3D video games.
+After every nine trials, the participants took a mandatory five minute break.
+Some experienced light fatigue, but none experienced strong effects of motion sickness.
 
@@
 In the target following state, the angular error is generally low.
-We asked participants to keep the cross hairs on the target as closely as they could, but different participants achieved different levels of accuracy.
+We had asked participants to keep the cross hairs on the target as closely as possible, but different participants achieved different levels of accuracy.
 In the lost state, the error starts to steadily climb, and may fluctuate depending on where the target moves.
@@
 
-We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.
-The threshold specifies the maximum angular error that still represents successful following of the target.
+We may detect the participant's state more explicitly by applying a threshold $\tau$ to the data.
+The threshold specifies the maximum angular error that represents successful following of the target.
 %When the error rises above the threshold, we consider the participant lost.
-While conducting the trials we noticed that the participants spent a majority of the time correctly following Top Hat, computing the median angular error at each time-step across all trials gave us values mostly around 4 degrees so we consider values close to this as successful tracking.
-We also noticed that during trials as Top Hat would change direction the participant's angular error would increase a small amount for a short time, without losing the target.
-As a result we doubled the 4 degrees, giving us a threshold of $\tau = 8$ degrees
+When conducting the trials we noticed that the participants spent a majority of the time correctly following Top Hat.
+By computing the median angular error, we found that the participants successfully tracked Top Hat within an error margin of four degrees.
+We also noticed that during trials as Top Hat would change direction the participant's angular error would increase by a small amount for a short period of time, without losing the target.
+As a result we doubled the four degrees, giving us a threshold of $\tau = 8$ degrees.
 % TODO: how to motivate this threshold?
@@
 	\includegraphics[width=1.5in]{figures/numtimes_fov.pdf}
-	\caption{\label{fig:tt_deadlen}Box plots for our two independent variables and two metrics.
+	\caption{\label{fig:boxplots}Box plots for our two independent variables and two metrics.
 In general, performance increases when the length of sensor dropout periods is lessened, and when the AR field of view increases.}
 \end{figure}
+
 Using this threshold we measure a participant's \emph{time to failure}, the time until the a lost state is encountered.
@@
 For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial, or  \emph{number of times lost}.
-We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 4 seconds.
-A 4 second threshold was used because it is double the maximum dropout length, allowing the user a few seconds to resume successful tracking after the dropout before they are considered to have lost the target.
+We only record a lost state when the error stays above the threshold $\tau$ for a minimum of four seconds.
+A four second threshold was used because it is double the maximum dropout length, allowing the user a few seconds to resume successful tracking after the dropout before they are considered to have lost the target.
 % TODO: how to motivate this threshold?
@@
 
+Figure \ref{fig:boxplots} shows box plots of two metrics versus our two independent variables..
 We performed two-way {ANOVA} for the time following target metric as an overall analysis of the effects of our two factors.
-Figure \ref{fig:anova} gives the results of this analysis.
-Both the failure length and the field of view have a significant effect on performance with this metric, at the $p=0.001$ level.
+%Figure \ref{fig:anova} gives the results of this analysis.
+Both the failure length and the field of view have a significant effect on performance with this metric with $p<0.001$.
 % TODO: interaction?
-
-We also used a two-way {ANOVA} to examine the number of times the participant loses the target, with results shown in Figure \ref{fig:anova}.
-Both field of view and dropout length have a significant effect at the $p=0.001$ level.
+We also used a two-way {ANOVA} to examine the number of times the participant loses the target.
+%, with results shown in Figure \ref{fig:anova}.
+Both field of view and dropout length have a significant effect with $p<0.001$.
 There also is an interaction between the two factors at the $p=0.05$ level.
-Figure \ref{fig:interaction} shows a plot of this interaction.
+Figures \ref{subfig:low}-(c) show box plots for these conditions.
 % TODO: need F values reported here.
@@
 
-\begin{figure}[t]
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=2.5in]{figures/numtimes_interaction.pdf}
+%	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
+%\end{figure}
+
+%\begin{figure*}[t]
+%	\centering
+%	\subfigure[]{\label{subfig:low}\includegraphics[width=1.5in]{figures/numtimes_deadlen_lowfov.pdf}}
+%	\subfigure[]{\label{subfig:med}\includegraphics[width=1.5in]{figures/numtimes_deadlen_medfov.pdf}}
+%	\subfigure[]{\label{subfig:upp}\includegraphics[width=1.5in]{figures/numtimes_deadlen_uppfov.pdf}}
+%	\caption{\label{fig:interaction}Box plots relating sensor dropout length to the number of times the participant loses the target, with (a) low, (b) medium and (c) upper field of view.}
+%\end{figure*}
+
+%\begin{figure}[h]
+%	\centering
+%	\includegraphics[width=1.5in]{figures/extra.pdf}
+%	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected this data from seven participants, using the medium field of view.}
+%\end{figure}
+
+\begin{figure*}[t]
 	\centering
-	\includegraphics[width=2.5in]{figures/numtimes_interaction.pdf}
-	\caption{\label{fig:interaction}Interaction plot of the effect of field of view and sensor dropout length on the number of times the participant loses the target.}
-\end{figure}
+	\subfigure[]{\label{subfig:low}\includegraphics[width=1.5in]{figures/numtimes_deadlen_lowfov.pdf}}
+	\subfigure[]{\label{subfig:med}\includegraphics[width=1.5in]{figures/numtimes_deadlen_medfov.pdf}}
+	\subfigure[]{\label{subfig:upp}\includegraphics[width=1.5in]{figures/numtimes_deadlen_uppfov.pdf}}
+	\subfigure[]{\label{subfig:extra}\includegraphics[width=1.5in]{figures/extra.pdf}}
+	\caption{\label{fig:interaction}Box plots (a)-(c) relate sensor dropout length to the number of times the participant loses the target, with (a) low, (b) medium and (c) upper field of view.
+Box plot (d) shows the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
+We collected the data for box plot (d) from seven participants, using the medium field of view.}
+\end{figure*}
 
-\begin{figure}[h]
-\centering
-\begin{tabular}{|c|c|c|c|}
-\hline
-%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
-Metric & Factor & F value & Pr \\
-\hline
-Time Following Target & Dropout Length & 28.904 & $<0.001$\\
-Time Following Target & Field of View & 15.800 & $<0.001$\\
-Number of Times Lost & Dropout Length & 8.5008 & $<0.001$\\
-Number of Times Lost & Field of View & 13.060 & $<0.001$\\
-\hline
-\end{tabular}
- \caption{\label{fig:anova}Factors which have a significant effect.}
-\end{figure}
-
-Figure \ref{fig:tt_deadlen} shows a box plot of time following target versus dropout length.
-Performance decreases between zero and one seconds, but seems to level off between one and two.
+%Figure \ref{fig:tt_deadlen} shows a box plot of time following target versus dropout length.
+Considering the time following target metric, performance decreases between zero and one seconds, but seems to level off between one and two.
 We ran a post-hoc analysis to determine which conditions were significantly different.
@@
 
+%\begin{figure}[h]
+%\centering
+%\begin{tabular}{|c|c|c|c|}
+%\hline
+%%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
+%Metric & Factor & F value & Pr \\
+%\hline
+%Time Following Target & Dropout Length & 28.904 & $<0.001$\\
+%Time Following Target & Field of View & 15.800 & $<0.001$\\
+%Number of Times Lost & Dropout Length & 8.5008 & $<0.001$\\
+%Number of Times Lost & Field of View & 13.060 & $<0.001$\\
+%\hline
+%\end{tabular}
+% \caption{\label{fig:anova}Factors which have a significant effect.}
+%\end{figure}
+
 To investigate this threshold, we had seven participants complete an extra three trials with the medium field of view and sensor dropout periods of 0.5 seconds.
-Figure \ref{fig:extra} shows a box plot of time following target versus sensor dropout length for these seven participants.
+Figure \ref{subfig:extra} shows a box plot of time following target versus sensor dropout length for these seven participants.
 Again using Tukey's HSD to compare the means, we did not find a significant difference between the zero length and half second length conditions, but did find a significant difference between those and the higher length conditions.
 %TODO: table with values
-With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one.
-
-\begin{figure}[htb]
-	\centering
-	\includegraphics[width=2.5in]{figures/extra.pdf}
-	\caption{\label{fig:extra}Box plot of the total time the participant is following the target versus the length of the sensor dropout periods, including the half second dropout condition.
-We collected this data from seven participants, using the medium field of view.}
-\end{figure}
+With further experimentation, we could further delineate the performance curve as the dropout length increases from zero to one second.
 
@@
 
-We found that for our task, good performance could not be achieved with a low field of view, even with perfect head-based rendering reliability.
+We found that for our target following task, good performance could not be achieved with a low field of view, even with perfect tracker reliability.
 %has an adverse effect on performance.   even when introducing medium-length sensor dropouts
 %This adverse effect dominates the other variable component of immersion, the reliability of the head-based rendering.
-With reliable head-based rendering, a medium field of view is enough for good performance.  We did not see a significant performance increase when increasing to a high field of view.
+With a reliable tracker, a medium field of view is enough for good performance.
+We did not see a significant performance increase when increasing to a high field of view.
+These results suggest that a reasonable AR field of view is crucial for performance in target following tasks, but the benefit does not continue to improve with increasing FOV.
 
-Decreased reliability in the head-based rendering, caused by periods of sensor dropout, adversely affects performance.
+Decreased tracker reliability, due to periods of sensor dropout, adversely affects task performance.
 However, the performance decrease levels off at one second dropouts.
 Preliminary analysis also shows no significant difference between having no dropouts and half second dropouts.
-This suggests that below a certain length of dropout time, good performance may be maintained, given a sufficiently large FOV.%  This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
+This suggests that below a certain dropout length, good performance may be maintained, given a sufficiently large FOV.
+%This threshold on dropouts would change depending on the task at hand, but we expect the same pattern would arise in other target following scenarios.
+Our findings suggest that AR interface designers should consider ways to mediate the effect of sensor dropouts longer than one second, although dropouts less than one half second long do not harm performance.
 
